### 2020-12-19 paper/literature_review.tex

\section{Literature Review}
\label{sec:literature_review}

We build on two strands of literature: Recent extensions of the epidemiological SEIR model and agent-based simulation models.

The traditional SEIR model is not fine-grained enough to model nuanced policies. This has motivated a large number of researchers to extend the standard model to allow for more heterogeneity and flexibility. Examples are \cite{Grimm2020}, \cite{Donsimoni2020} and \cite{Acemoglu2020} who develop multi group SEIR models to analyze the effects of targeted lockdowns and \cite{Berger2020} who extend the SEIR model to analyze testing and conditional quarantines. For a more comprehensive review see \cite{Avery2020}. Others have used the results of a standard SEIR model as input for economic models that estimate the cost of policies (e.g. \cite{Dorn2020}).

While the popularity of the SEIR model is mainly due to its simplicity, the extensions are quite complex. It is unlikely that there will be a SEIR model that combines all proposed extensions. Moreover, the extensions do not address other key issues: The main parameter of the SEIR model, the basic reproduction number ($R_0$), is not policy-invariant. It is a composite of the number of contacts each person has and the infection probability of the contacts. In fact, policy simulations are done by setting $R_0$ to a different value but it is hard to translate a real policy into the value of $R_0$ it will induce.

Another commonly used model class in epidemiology are agent-based simulation models. In these models individuals are simulated as moving particles. Infections take place when two particles come closer than a certain contact radius (e.g. \cite{Silva2020} and \cite{Cuevas2020}). While the simulation approach makes it easy to incorporate heterogeneity in disease progression, it is hard to incorporate heterogeneity in meeting patterns. Moreover, policies are modeled as changes in the contact radius or momentum equation of the particles. The translation from real policies to corresponding model parameters is a hard task.

\cite{Hinch2020} is a recent extension of the prototypical agent-based simulation model that replaces moving particles by contact networks for households, work and random contacts. This model is similar in spirit to ours but focuses on contact tracing rather than social distancing policies.

The above assessment of epidemiological models is not meant as a critique. We are aware that those models were not designed to predict the effect of fine grained social distancing policies in real time and are very well suited to their purpose. We invite epidemiologists to provide feedback and collaborate to improve our model.