### 2020-12-19 paper/model.tex
Split up the section on the model in subfiles, added a file on testing.

\section{Model}
\label{sec:model}

\input{model/summary}

\input{model/number_of_contacts}

\input{model/policies}

\input{model/endogenous_contact_reductions}

\input{model/matching}

\FloatBarrier

\input{model/course_of_disease}

\input{model/testing}