### 2020-12-19 paper/model/endogenous_contact_reductions.tex

\subsection{Endogenous Contact Reductions}
\label{sec:endogenous_contact_reductions}

Policies are not the only way in which the number of contacts are reduced compared to the pre-pandemic level. It is important to model those other channels. Otherwise, the effect of policies would be overestimated and policy recommendations based on the model would be biased.

Examples of endogenous contact reductions are manifold: symptomatic people stay at home; Members of risk groups try to reduce their number of contacts more strongly than others; People self-isolate if they know they had a risk contact.

Since we model the number of contacts as arbitrary functions of background characteristics and health states, it is easy to implement such considerations.

In our current empirical application we only model that symptomatic people reduce their number of contacts across all contact types (except for households) by 70 \%. Within households they reduce contacts by 50\%. We are working on extending this to allow for formal and informal contact tracing as well as quarantines after positive test results.