### 2020-12-19 paper/conclusion.tex
Clean up preamble, fix typos and add section on initial conditions.

\section{Conclusion}
\label{sec:conclusion}

We propose a simulation based model of infectious disease transmission that is designed to predict the effects of fine-grained social distancing policies. In particular, the model can be used to model policies such as several ways of splitting school classes or work reduction policies that affect essential and non-essential workers differently. Both policies would be hard to implement in standard SEIR or agent based simulation models.

To predict the effects of such policies, it is not only important to have a way of expressing such flexible policies in terms of model quantities, but also to incorporate heterogeneity in disease progression as well as meeting patterns. We calibrate age dependent disease progression parameters from the medical literature and age dependent contact frequencies from contact diaries. Moreover, we distinguish ten types of contacts out of which some are only relevant for certain age groups.

The model has a good fit on past German case numbers and passes an out of sample validation despite a drastic change in the policy environment between the estimation period and the validation period.

Despite these encouraging results we still see the model as work in progress and plan to implement more features such as a detailed model of testing and contact tracing. Moreover, the estimation of the infection probabilities and the model fit will improve as more data becomes available.

We invite researchers from any discipline, but particularly epidemiologists to provide feedback on the model and welcome collaborations.