### 2020-12-19 paper/model/initial_conditions.tex

\subsection{Initial Conditions} % (fold)
\label{sub:initial_conditions}

Consider a situation where you want to start a simulation with the beginning set amidst the pandemic. It means that several thousands of individuals have already recovered from the disease, are infectious, symptomatic or in ICUs. Additionally, the sample of infectious people who will determine the course of the pandemic in the following periods is likely not representative of the whole population because of differences in behavior (number of contacts, assortativeness), past policies (school closures), etc.. The distribution of courses of diseases in the population at the begin of the simulation are called initial conditions.

To come up with realistic initial conditions, we match reported infections from official data to simulated individuals by available characteristics like age and geographic information. The matching must be done for each day of a longer time frame like a month to have individuals in all possible health states. Then, health statuses evolve until the begin of the simulation period without simulating infections by contacts. It is also possible to correct reported infections for a reporting lag and scale them up to arrive at the true number of infections.

% subsection initial_conditions (end)