### 2020-12-20 paper/model/course_of_disease.tex
Enforced 88 character limit.

\subsection{Course of the Disease}
\label{sub:course_of_disease}

The following medical parameters describing the progression of the disease are taken
from systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
progresses in the way depicted in Figure~\ref{fig:course_of_disease}.

\begin{figure}[!tp]
    \includegraphics[width=\textwidth]{../figures/disease_progression.png}
    \caption{Course of Disease in the model}
    \label{fig:course_of_disease}
\end{figure}

First, infected individuals will become infectious after one to five days. About one
third of people stays asymptomatic. The rest develops symptoms about one to two days
after they become infectious. Modeling asymptomatic and pre-symptomatic cases is
important because those people do not reduce their contacts or demand a test and can
potentially infect many other people \citep{Donsimoni2020}.

A small share of symptomatic people will develop strong symptoms that require intensive
care. The exact share and time span is age-dependent. An age-dependent share of
intensive care unit (ICU) patients will die after spending up to 32 days in intensive
care. Moreover, if the ICU capacity was reached, all patients who require intensive care
but do not receive it die.

It would be easy to make the course of disease even more fine-grained. For example, we
could model people who require hospitalization but not intensive care. So far we opted
against that because only the ICU capacity might become a bottleneck in Germany.

We allow the progression of the disease to be stochastic in two ways: Firstly, state
changes only occur with a certain probability (e.g. only a fraction of infected
individuals develops symptoms). Secondly, the number of periods for which an individual
remains in a state is drawn randomly. The parameters that govern these processes are
taken from the literature. They can vary with the age of an individual.

Detailed information on the calibration of the disease parameters is available as part
of our
\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
documentation}.