### 2021-02-08 paper/model/summary.tex

\subsection{Summary}
\label{sub:model_summary}

To predict the effects of a wide variety of fine-grained social distancing policies
we propose a different model structure.
Our model inherits many features from agent-based simulation models but replaces
the contacts between moving particles by contacts between individuals who work,
go to school, live in a household and enjoy leisure activities.
The structure of the model is depicted in Figure~\ref{fig:model_graph}.

\begin{figure}[!tp]
    \centering
    \includegraphics[width=\textwidth]{../figures/model_detailed.png}
    \caption{Simplified graph of the model.}
    \label{fig:model_graph}
\end{figure}

The background characteristics include age, county and occupation of each simulated
individual. Contact models are functions that map individual characteristics into a
predicted number of contacts. Currently, we distinguish between eight types of contact
models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and
random work contacts, recurrent and random leisure contacts, and nursery, preschool, and
school contacts.

The predicted number of contacts is translated into infections by a matching algorithm.
There are different matching algorithms for recurrent contacts (e.g. classmates, family
members) and non-recurrent contacts (e.g. clients, contacts in supermarkets). The
infection probability can differ for each contact type. All types of contacts can be
assortative with respect to geographic and demographic characteristics.

Once a person is infected, the disease progresses in a fairly standard way which is
depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild
symptoms are infectious for some time and recover eventually. Cases with severe symptoms
additionally require hospitalization and lead to either recovery or death.

Our model also allows the evaluation of different testing strategies by having a detailed
three-step model of test demand, allocation and processing:
First, individuals demand a test because they, e.g., experience symptoms.
Secondly, tests are allocated to individuals while respecting governmental access
restrictions to test.
Thirdly, depending of the capacities of laboratories, tests are processed for some time
until the individual receives her test result.

In addition, people who have symptoms, received a positive test, or had a risk contact
can reduce their number of contacts across all contact types endogenously.

The model makes it very simple to translate policies into model quantities. For example,
school closures imply the complete suspension of school contacts. A strict lockdown
implies shutting down work contacts of all people who are not employed in a systemically
relevant sector. It is also possible to have more sophisticated policies that condition
the number of contacts on observable characteristics, risk contacts or health states.

Another key advantage of the model is that the number of contacts an individual has of
each contact type can be calibrated from publicly available data \citep{Mossong2008}.
This in turn allows us to estimate policy-invariant infection probabilities from time
series of infection and death rates using the method of simulated moments
\citep{McFadden1989}. Since the infection probabilities are time-invariant, data
collected since the beginning of the pandemic can be used for estimation. Moreover,
since we model the testing strategies that were in place at each point in time, we
can correct the estimates for the fact that not all infections are observed.

Last but not least, performing simulations whose starting point is set amidst the
pandemic requires special adjustments to arrive at a realistic distribution of courses
of diseases. We solve the initial conditions problem by matching reported infections to
individuals in our data while also correcting for reporting lag and undetected cases.

In the following sections we describe each of the model components in more detail.