### 2021-02-09 paper/calibration_and_estimation.tex
Update calibration and estimation section.

\section{Calibration and Estimation}
\label{sec:calibration_and_estimation}

The model is described by a large number of parameters that govern the number of
contacts a person has, the likelihood of becoming infected on each contact, the
likelihood of developing light or strong symptoms or even dying from the disease as well
as the duration each stage of the disease takes.

Most of these parameters can be calibrated from existing datasets or the medical
literature. Only the infection probabilities have to be estimated inside the model by
fitting it to time series data of case numbers and fatality rates.

\input{medical_parameters}

\subsection{Number of Contacts}
\label{sub:number_of_contacts}

We calibrate the parameters for the predicted numbers of contacts from contact diaries
of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
\citep{Mossong2008}. Each contact diary contains all contacts an individual had
throughout one day, including information on the other person (such as age and gender)
and information on the contact. Importantly, for each contact individuals entered of
which type the contact (school, leisure, work etc.) was and how frequent the contact
with the other person is.

Thus, we can use the empirical distributions from this data as pre-pandemic number of
contacts.

\FloatBarrier

\subsection{Assortative Matching}

As mentioned in section \ref{sec:matching}, the probability that two individuals are
matched can depend on background characteristics. In particular, we allow this
probability to depend on age and county of residence. While we do not have good data on
geographical assortativeness and just roughly calibrate it such that 60 \% of contacts
are within the same county, we can calibrate the age assortativeness from
the same data we use to calibrate the number of contacts.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9 \textwidth]{../figures/assortative_matching_probability_example.png}
    \caption{Distribution of random non-work contacts by age of participants.}
    \label{fig:assortativeness}
    \figurenotes{
        The figure shows the distribution of random non-work contacts by age groups. A
        row shows the share of contacts a certain age group has with all other age
        groups. Higher values are colored in darker red tones. The diagonal represents
        the share of contacts with individuals from the same age group.
    }
\end{figure}

Figure~\ref{fig:assortativeness} shows that assortativeness by age is especially strong
for children and younger adults. For older people, the pattern becomes more dispersed
around their own age group, but within-age-group contacts are still the most common
contacts.

\FloatBarrier

\subsection{Infection Probabilities}
\label{sec:estimation}

To calibrate infection probabilities outside of the model, it would be important to know
the exact duration and distance of each contact type as well as viral loads. Since this
is not available in any dataset, we estimate those parameters inside the model with the
method of simulated moments \citep{McFadden1989} by minimizing the distance between
simulated and observed infection rates. Since our model includes a lot of randomness, we
average simulated infection rates over several model runs.

Currently, we use data for Germany from August until November.
We do not use earlier periods to save computational time.
Moreover, we would be worried that the there are seasonal effects that we currently do
not model.

To avoid overfitting and simplify the numerical optimization problem, we only allow for
four different probabilities: 1) for contacts in schools, preschools and nurseries. 2)
for work contacts. 3) for households. 4) for leisure activities.

\subsection{Policies}

\FloatBarrier

In our empirical application we distinguish four groups of contact types:
households, education, work and other contacts.
For households we assume that the individuals' contacts in their households do not change
over our estimation period.
For nurseries, preschools and schools we implement vacations as announced by the German
federal states as well as school closures. For the moment we ignore both emergency
childcare and that lack of childcare leads working parents to stay home.
%
% Schließung von Kindertagesstätten und Schulen: 37,4 Millionen ausgefallene Arbeitstage
% http://www.iab-forum.de/schul-und-kitaschliessungen-krankheit-quarantaene-die-coronabedingten-arbeitsausfaelle-der-erwerbstaetigen-steigen-auf-592-millionen-arbeitstage/
%
%
% https://www.sueddeutsche.de/politik/schulschliessung-lockdown-bildung-1.5190377:
% In allen Ländern geht trotz des Lockdowns ein erheblicher Anteil der Schülerinnen und
% Schüler in die Schule.
% https://gfx.sueddeutsche.de/apps/e525337/www/_image_desktopw1840q70-1e2e2bf78b7d4430.png
% 18% der Grundschüler in Notbetreuung in BW
%
For our work models\footnotemark we use the reductions in work mobility reported in the
Google Mobility Data \citep{Google2021} to calibrate our work policies.
Reductions in work contacts are not random but governed through a work contact priority
where the policy changes the threshold below which workers stay home.
Figure \ref{fig:work_multiplier} shows the share of workers that go to work in our model
over time.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9 \textwidth]{../figures/work_multiplier.png}
    \caption{Share of workers attending work normally.}
    \label{fig:work_multiplier}
    \figurenotes{
        The figure shows the percentage of pre-pandemic work mobility taking place
        since the start of the pandemic. In our model this is used as a proxy for the
        share of workers that attend work as usual. Workers are ordered with the
        importance of personal contacts for their work. Thus, if the share of
    }
\end{figure}

\footnotetext{We distinguish non-recurrent work contacts, daily work contacts and weekly work contacts.}

For the last group of contacts which cover things like leisure activities, grocery
shopping etc. we have no reliable data by how much policies reduce them.
In addition, they are likely to be affected by social and psychological factors such as
pandemic fatigue and vacations. Because of this we estimate them like the infection
probabilities to fit the time series data. We use very few change points and tie them to
particular events such as policy announcements or particular holidays.

\FloatBarrier