### 2021-04-17 paper/abstract.tex
Bring paper up to date for CRC Discussion Series. (#2)
Co-authored-by: Tobias Raabe <raabe@posteo.de>

\noindent
Governments worldwide have been adopting diverse and nuanced policy measures to contain
the spread of Covid-19. However, epidemiological models usually lack the detailed
representation of human meeting patterns to credibly predict the effects such policies.
%
We propose a novel simulation-based model to address these shortcomings. We build on
state-of-the-art agent-based simulation models, greatly increasing the amount of detail
and realism with which contacts take place. Firstly, we allow for different contact
types (such as work, school, households or leisure), distinguish recurrent and
non-recurrent contacts and allow the infectiousness of meetings to vary between contact
types.
% Secondly, we strictly separate the number of contacts from the probabilities that a
% contact leads to an infection.
Secondly, we allow agents to seek tests and react to information, such as experiencing
symptoms, receiving a positive test or a known case among their contacts, by reducing
their own contacts.
% Thus, social distancing policies can affect the number of contacts while the infection
% probabilities remain invariant.
This allows us to model the effects of a wide array very targeted policies such as split
classes, mandatory work from home schemes or test-and-trace policies. To validate our
model, we show that it can predict the effect of the German November lockdown even if no
similar policy has been observed during the time period that was used to estimate the
model parameters.

\vspace{1cm}
\noindent JEL Classification: C63, I18

\noindent Keywords: Covid-19, agent based simulation model, public health measures