### 2021-04-17 paper/model/testing.tex

\subsection{Testing} % (fold)
\label{sub:testing}

We support to model testing as consisting of three stages. Firstly, we model who demands
a test. Demand functions map from individual characteristics to a probability which is
the probability for this individual to demand a test. There can be multiple demand
functions where each function may describe a different channel. For example, individuals
who experience symptoms or have a risk contact may ask for a test. Or, the ministry of
education requires a negative test result from every teacher every second week. After
the probabilities for each individual and every demand model are computed, individuals
who demand a test as well as the channel is sampled.

The second stage is the allocation phase in which demand and supply for tests are
matched. The number of available tests can be inferred from official data and used to
model shortages in supply. When demand exceeds supply, some individuals might be given
preferred access to tests because of their own vulnerability or their potential to
become a super-spreader.

In the last and third phase, administered tests are processed. This step can become a
bottleneck in the testing process if there are not enough laboratories or necessary
resources available to evaluate all tests.

In our empirical estimation we use a very simplified testing model where the number of
tests to be distributed is calculated from estimates for the ratio of known to all
infections.\footnote{The Dunkelzifferradar project publishes daily estimates of the dark
figure of infections under \url{https://covid19.dunkelzifferradar.de/}} Using these
estimates as well as data on the test distribution over age groups by the
RKI\footnote{https://ars.rki.de/Content/COVID19/Main.aspx} we allocate tests firstly
among the symptomatic and then randomly allocate tests to newly infected to fit the
German test distribution.

% subsection testing (end)