### 2021-06-01 paper/results.tex
Add results section to paper and results.

\section{Results}
\label{sec:results}

This is the results section.

\textcolor{red}{The results can be found in `figures/results`. This includes both figures
and tables for lookup of numbers and summary tables}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/comparisons/one_off_and_combined/full_new_known_case}
    \caption{The Effect of Each Channel on Observed New Cases Separately}
    \label{fig:}
    \figurenotes{\textcolor{red}{\ldots}}
\end{figure}