### 2021-06-09 paper/medical_parameters.tex
Fix reference unicode error. Minor improvements to the medical parameters.

\subsection{Medical Parameters}
\label{sec:medical_params}

This section discusses the medical parameters used in the model, their sources and how we
arrived at the distributions used in the model.\footnotemark See
Figure~\ref{fig:disease_progression} for a summary of our disease progression model.

\footnotetext{Additional information can be found in the
\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
documentation}.}

\subsubsection{Duration until Infectiousness and Symptoms}

% Latency Period
The first medical parameter we need is the length of the period between infection and the
start of infectiousness, the so called latent period. We infer it from two other measures
that are more common in the medical literature: Firstly, the time between infection and
the onset of symptoms, the incubation period. Secondly, the time between the start of
infectiousness and the onset of symptoms. We assume that the latency period is the same
for symptomatic and asymptomatic individuals.

% Asymptomatic Share
Once individuals become infectious a share of them goes on to develop symptoms while
others remain asymptomatic. We rely on data by \cite{Davies2020} for the age-dependent
probability to develop symptoms. It varies from 25\% for children and young adults to
nearly 70\% for the elderly.

% Incubation Period
The incubation period is usually estimated to be two to twelve days. A meta analysis by
\citet{McAloon2020} comes to the conclusion that ``The incubation period distribution may
be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\%
CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For
simplicity we discretize this distribution into four bins.

% Length of the presymptomatic stage
The \href{https://bit.ly/3pi18Ax}{European Centre for Disease Prevention and Control}
reports that people become infectious between one and two days before symptoms
start.\footnotemark

\footnotetext{This is similar to \citet{He2020} and in line with \citet{Peak2020}.}

Taking these estimates together, we arrive at a latent period of one to five days.

\subsubsection{Duration of Infectiousness}

We assume that the duration of infectiousness is the same for both symptomatic and
asymptomatic individuals as evidence suggests little differences in the transmission
rates between symptomatic and asymptomatic patients
(\citet{Yin2020}) and that the viral load between symptomatic and asymptomatic
individuals are similar (\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}).

Our distribution of the duration of infectiousness is based on \citet{Byrne2020}.

For symptomatic cases they arrive at zero to five days before symptom onset (figure 2)
and three to eight days of infectiousness afterwards.\footnotemark Thus, we arrive at 0
to 13 days as the range for infectiousness among individuals who become symptomatic (see
also figure 5).

\footnotetext{Viral loads may be detected much later but eight days seems to be the time
    after which most people are culture negative, as also reported by
    \citet{Singanayagam2020}.}

\subsubsection{Duration of Symptoms}

We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3,
Panel~2]{Bi2020} for the duration of symptoms for asymptomatic and non-ICU requiring
symptomatic cases.

These numbers are only used for mild cases. We do not disaggregate by age. Note that the
length of symptoms is not very important in our model given that individuals stop being
infectious before their symptoms cease.

\subsubsection{Time from Symptom Onset to Admission to ICU}

The data on how many percent of symptomatic patients will require ICU is pretty thin. We
rely on data by the US CDC (\citet{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the
OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the
probabilities of requiring intensive care per age group.

\begin{table}[tb]
    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
    \label{tab:symptomatic-to-ICU}
    \centering

    \begin{tabular}{ll}
        \toprule
        Age Group & Share \\
        \midrule
        0-9 & 0.00005 \\
        10-19 & 0.00030 \\
        20-29 & 0.00075 \\
        30-39 & 0.00345 \\
        40-49 & 0.01380 \\
        50-59 & 0.03404 \\
        60-69 & 0.10138 \\
        70-79 & 0.16891 \\
        80-100 & 0.26871 \\
        \bottomrule
    \end{tabular}

    \tablenotes{
        The data is taken from \citet{Stokes2020} and
        \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the
        OpenABM-Project}. }

\end{table}

For those who will require intensive care we follow \citet{Chen2020} who estimate the
time from symptom onset to ICU admission as 8.5 $\pm$ 4 days.

This aligns well with numbers reported for the time from first symptoms to
hospitalization: \citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation
of 4. This is also in line with the durations collected by
\href{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Steckbrief.html#doc13776792bodyText16}{the
Robert-Koch-Institut}.

We assume that the time between symptom onset and ICU takes 4, 6, 8 or 10 days with
equal probabilities. These times mostly matter for the ICU capacities.

\subsubsection{Death and Recovery from ICU}

We take the survival probabilities and time to death and time until recovery from
intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}.

They report time until death to have a mean of 11.74 days and a standard deviation of
8.79 days. Approximating this with the normal distribution, we have nearly 10\%
probability mass below 0. We use it nevertheless as several other distributions (such as
chi squared and uniform) were unable to match the variance. Discretizing the
distribution leads to 41\% of individuals who will die from Covid-19 after one day in
intensive care, 22\% day after 12 days, 29\% after 20 days and 7\% after 32 days. Again,
we rescale this for every age group among those that will not survive.

They report a mean duration of 18.8 days until recovery and a standard deviation of
12.21 days. Approximating this with the normal distribution, we have over 5\%
probability mass below 0. Of those who recover in intensive care, 22\% do so
after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.