### 2021-06-09 paper/appendix_model.tex
Re-add literature review.

\section{Detailed Model Description}
\label{sec:model}

\input{literature_review}

\input{model/summary}

\input{model/number_of_contacts}

\input{model/policies}

\input{model/endogenous_contact_reductions}

\input{model/matching}

\FloatBarrier

\input{model/course_of_disease}

\input{model/testing}

\input{model/initial_conditions}