### 2021-06-10 paper/appendix_model.tex
Work on first batch of comments in report.tex.

\section{Detailed Model Description}
\label{sec:model}

\input{literature_review}

\input{model/summary}

\input{model/number_of_contacts}

\input{model/policies}

\input{model/endogenous_contact_reductions}

\input{model/matching}

\FloatBarrier

\input{model/course_of_disease}

\input{model/testing}

\input{model/initial_conditions}

\textcolor{red}{Further things:}

\begin{itemize}
    \item talk more in detail about vaccinations? (35\% reach immunity after two weeks, 40\% after three weeks, distribution with the STIKO groups and children and youths...)
\end{itemize}