### 2021-06-11 paper/appendix_results.tex
Add rapid test statistics figures. Add share known cases and rapid test statistics to the appendix.

\section{Additional Results}
\label{sec:additional_results}

\subsection{Simulated vs. Empirical Data}
\label{subsec:new_cases_fit}

This compares simulated data from our model with empirical data from Germany. We look at
observed infections, fatality rates, the spread of the B117 mutation, vaccinations and
rapid test demands. Where available we do not only look at aggregated statistics but also
analyze the model fit for age groups.\comment[id=J]{summarize the fit; Some of this needs
to go in the appendix. I suggest the age group and federal state fit goes there???}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case_with_single_runs}
  \caption{Simulated and Empirical Infections}
  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates.}
  \label{fig:aggregated_fit2}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/age_group_rki/full_combined_baseline_new_known_case}
  \caption{Simulated and Empirical Infections by Age Group}
  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different age groups.}
  \label{fig:age_group_fit}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../figures/results/figures/incidences_by_group/state/full_combined_baseline_new_known_case}
  \caption{Simulated and Empirical Infections by Federal State}
  \figurenotes{The figure shows the weekly incidence rates per 100,000 people for the reported versus the simulated infections rates for different federal states.}
  \label{fig:state_fit}
\end{figure}

\FloatBarrier

\subsection{Share of Cases that are Detected}
\label{subsec:appendix_share_known_cases}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../figures/results/figures/share_known_cases/full_combined_baseline_by_age_group_rki}
  \caption{Share of Detected Cases by Age Group in the Main Prediction}
  \label{fig:share_known_cases_by_age_group}
  \floatfoot{\noindent \textcolor{red}{\ldots}}
\end{figure}

It's noteworthy that the share of detected cases increases rapidly in May for the five to fourteen year olds. This is a direct result of the mandatory tests in school.\comment[id=K]{We don't do this at the moment but we could also create plots showing the share of detected cases between scenarios - either for the full population or for a age group of interest.}

\subsection{Rapid Tests}
\label{subsec:appendix_rapid_tests}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/demand_shares}
  \caption{Share of the Population Demanding a Rapid Test for a Particular Reason}
  \label{fig:rapid_tests_by_reason}
  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without PCR demand and other contact demand together form the private demand category. Also note that these do not add up to the total share of demanded rapid tests as individuals may have more than one reason to demand a rapid test on any given day. \textcolor{red}{\ldots}}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../figures/results/figures/rapid_test_statistics/share_infected_among_demand}
  \caption{Share of Individuals that Demand a Rapid Test for a Particular Reason that are Actually Infected}
  \label{fig:share_of_rapid_tests_by_reason_by_infected}
  \floatfoot{\noindent Note that the lower three (household demand, symptomatic without PCR demand and other contact demand together form the private demand category. Also note that this is not the same as individuals getting a positive rapid test. The sensitivity is quite low before individuals become infectious. Therefore, if individuals are still in the latent period of their infection they are likely to get a false positive rapid test. \textcolor{red}{\ldots}}
\end{figure}

\FloatBarrier

\subsection{Scenarios}
\label{subsec:appendix_scenarios}

This is the results section.\comment[id=K]{The results can be found in `figures/results`.
This includes both figures and tables for lookup of numbers and summary tables.}

\begin{figure}[ht]
  \centering
  \begin{subfigure}{.6\textwidth}
    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_new_known_case_cropped}
  \end{subfigure}%
  \begin{subfigure}{.6\textwidth}
    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/one_off_and_combined/full_newly_infected_cropped}
  \end{subfigure}
  \caption{The Effect of Policies on Observed and Unobserved Cases}
  \label{fig:explain_decline}
  \figurenotes{\ldots}
\end{figure}

\begin{figure}[ht]
  \centering
  \begin{subfigure}{.6\textwidth}
    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_new_known_case}
  \end{subfigure}%
  \begin{subfigure}{.6\textwidth}
    \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
  \end{subfigure}
  \caption{The Effect of Different School Scenarios on Observed and Unobserved Cases}
  \label{fig:school_scenarios_detailed}
\end{figure}

\FloatBarrier

\input{../figures/results/tables/student_infections_table}

\FloatBarrier