### 2021-06-14 paper/report.tex
Add Janos' comments on the report.

% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3
% printed pages in the journal) present important new research results of broad
% significance. Reports should include an abstract, an introductory paragraph, up to
% four figures or tables, and about 30 references. Materials and Methods should be
% included in supplementary materials, which should also include information needed to
% support the paper's conclusions.

% Main Text is not divided into sub-headings for Reports.

% The manuscript should start with a brief introduction describing the paper’s
% significance. The introduction should provide sufficient background information to
% make the article intelligible to readers in other disciplines, and sufficient context
% that the significance of the experimental findings is clear.

Since early 2020,\comment[id=J]{\textbf{Features that are not highlighted enough:}

\begin{itemize}
    \item Low sensitivity of rapid tests before and on day of infectiousness onset
    \item Different seasonalities for different contact types (and quite strong for other contacts)
    \item Low number of breakpoints in parameters / policies compared to breakpoints in
    cases / reproduction rate
\end{itemize}

} the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many
months---and often years---away from reaching vaccination-induced herd immunity
\citep{Swaminathan2021}.\comment[id=HM]{Need to find some more scientific citation
before submission} In the meantime, it is of utmost importance to employ an effective
mix of strategies for containing the virus. The most frequent initial response was a set
of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals. While
this has allowed some countries to sustain equilibria with very low infection
numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case
numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
fluctuations of infection rates over time. Containment measures have become increasingly
diverse and now include testing, more nuanced NPIs, and contact tracing. Neither these
policies' effect nor the influence of seasonal patterns or more infectious virus strains
are well understood in quantitative terms.

This paper develops a quantitative model incorporating all these factors. The framework
allows to combine a wide variety of data and mechanisms in a timely fashion, making it
useful to predict the effects of various interventions. We apply the model to Germany
and show that rapid testing had the largest impact on the reduction in infections by
almost 80\% during the month of May 2021. We conclude that rapid tests have a large role
to play for at least as long as vaccinations have not been offered to an entire
population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:broad_model_description}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
model most closely related to ours is described in \citet{Hinch2020}.} Each contact
between an infectious individual and somebody susceptible to the disease bears the risk
of transmitting the virus. Contacts occur in the household, at work, at school, or in
other settings (leisure activities, grocery shopping, medical appointments, etc.). Some
contacts recur regularly, others occur at random. Random contacts are typically
assortative in age and geographical location. Empirical applications can take the
population structure from census data and the types and frequency of contacts from diary
data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008,Hoang2019}.

The three dimensions of contact types are chosen so that the most common NPIs can be
modeled in great detail. NPIs affect the number of contacts or the risk of transmitting
the disease upon having physical contact. The effect of different NPIs will generally
vary across contact types. For example, a mandate to work from home will reduce the
number of work contacts to zero for a fraction of the working population. Schools and
daycare can be closed entirely, operate at reduced capacity---including an alternating
schedule---,or implement mitigation measures like masking requirements or air filters
\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
settings. In any setting, measures like masking requirements would reduce the
probability of infection associated with a contact \citep{Cheng2021}.

\begin{figure}
    \centering

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
        \caption{{Model description}}
        \label{fig:broad_model_description}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
        \vskip5ex

        \caption{Disease progression}
        \label{fig:disease_progression}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering

        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
        \caption{{Model for antigen tests}}
        \label{fig:antigen_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
        \vskip5ex

        \caption{{Translating all infections to recorded ones}}
        \label{fig:model_for_official_cases}
    \end{subfigure}

    \caption{Model description}
    \label{fig:model-description}

    \floatfoot{\noindent \textit{Note:}
    A % formal 
    description of the model can be found in Supplementary Material~\ref{sec:model}.
    Figure~\ref{fig:broad_model_description} shows the influence on an agent's contacts
    to other agents. Demographic characteristics set the baseline number of contacts in
    different networks. She may reduce the number of contacts due to NPIs, showing
    symptoms, or testing positively for SARS-CoV-2. Infections may occur when a
    susceptible agent meets an infectious agent; the probability depends on the type of
    contact, on seasonality, and on NPIs. If infected, the infection progresses as
    depicted in Figure~\ref{fig:disease_progression}. If rapid tests are avaible,
    agents' demand is modeled as in Figure~\ref{fig:antigen_tests}. All pupils are
    attending school are tested, other reasons trigger a test only for a fraction of
    individuals. Figure~\ref{fig:model_for_official_cases} shows the model of
    translating all infections in the simulated data to age-specific recorded
    infections. The model uses data on the aggregate share of recorded cases, on the
    share of positive PCR tests triggered by symptoms, and on the false positive rate of
    rapid tests. The lower part of the graph is relevant only for periods where rapid
    tests are available.}
\end{figure}

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
The probabilities of transitioning between these states depend on age; their duration is
random within intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.

The model includes several other features, which are crucial to describe the evolution
of the pandemic in 2020-2021. New virus strains with different profiles regarding
infectiousness can be introduced. Agents may receive a vaccination. With a probability
of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not transmit the
virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\% is lower
    than what is usually reported for after the second dose of the Biontech/Pfizer vaccine,
    which is most commonly used in Germany. We choose it because our model neither includes
    booster shots, nor does it allow vaccinated individuals who became immune to transmit
    the disease\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}. If anything, these
    assumptions would overstate the effect of vaccines for our study period. This would be
    different if a large fraction of vaccinated individuals had received a second dose
    already.} During the vaccine roll-out, priority may depend on age and occupation.

We include two types of tests. Polymerase chain reaction (PCR) tests
directly\comment[id=J]{It is not clear what "directly" means} reveal whether an
individual is infected or not. PCR tests require some days to be processed and there are
aggregate capacity constraints throughout. In contrast, rapid antigen tests yield
immediate results; after a phase-in period, all tests that are demanded will be
performed. Specificity and sensitivity of these tests is set according to data analyzed
in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test relative
to the start of infectiousness. Figure~\ref{fig:antigen_tests} shows our model for rapid
test demand. Schools may require students to be tested regularly. Rapid tests may be
offered by employers for on-site workers. Individuals may demand tests for private
reasons, which include having plans to meet other people\footnote{A positive test will
make them reduce their contacts; this is why tests impact the actual contacts in
Figure~\ref{fig:model-description}.}, showing symptoms of CoViD-19, and because a
household member tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid tests offered
by employers or follows up on private reasons and whether she reduces her contacts if she
tests positive or develops symptoms. The thresholds are higher for tests in a private
setting than for tests at the workplace until mid April and then compliance because of
private reasons becomes higher than compliance at the workplace.\comment[id=J]{Shouldn't
we mention that this is because of availability / high costs?}

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group. Mobility data is used
to model the reductions in work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests.

The free parameters are the infection probabilities by contact type, see Supplementary
Material~\ref{subsec:infection_probs}. We estimate these parameters so that simulated
infections by age predict the corresponding time series in the data \citep[See][for the
general method]{McFadden1989}. However, the two sets of time series are not comparable
directly because not all cases will be recorded. We thus model recorded cases as depicted
in Figure~\ref{fig:model_for_official_cases}. We take aggregate estimates of the share of
recorded cases and use data on whether the reason for a PCR test were symptoms of
CoViD-19 to get age-specific shares. For example,\comment[id=HM]{@Janos, please correct
if necessary and insert reference} estimate unknown cases to be twice as high for
children as for adults. Confirmation of a positive rapid test result is another reason
for PCR tests, once these become available.\comment[id=J]{Free parameters also include:

- One hygiene multiplier for school an work contacts after November 2
- A few "other" multipliers including two lockdown fatigue effects
- Number of extra contacts during holidays

Strictly speaking also:
- The fade in speed of the rapid tests
- Share known cases around holidays }

The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the Oxford Response
Stringency Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical
interventions. We transform the index so that lower values represent higher levels of
restrictions. A value of zero means all measures incorporated in the index are turned
on. The value 1 represents the situation in mid-September, with restrictions on
gatherings and public events, masking requirements, but open schools and workplaces. In
the seven weeks between mid September and early November, cases increased by a factor of
10. Restrictions were somewhat tightened in mid-October and again in early November. New
infections remained constant throughout November, before rising again in December, which
prompted the most stringent lockdown to this date. Schools and daycare centers were
closed again, so were customer-facing businesses except for grocery and drug stores.
From the peak of the second wave just before Christmas until the trough in mid-February,
newly detected cases decreased by almost three quarters. The third wave in the spring of
2021 is associated with the B.1.1.7 strain, which became dominant in March. See
Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
stringency index.

\begin{figure}[!tp]
    \centering

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
        \caption{{Recorded cases: Empirical and simulated}}
        \label{fig:aggregated_fit}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/results/figures/data/stringency2_with_seasonality}

        \caption{{Stringency of NPIs and infectious contacts}}
        \label{fig:stringency_infectious_contacts}
    \end{subfigure}

    \vskip3ex

    \begin{subfigure}[b]{0.475\textwidth}
        \centering

        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}

        \caption{{Tests and vaccinations}}
        \label{fig:antigen_tests_vaccinations}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering

        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}

        \caption{Fraction of B.1.1.7 strain}
        \label{fig:share_b117}
    \end{subfigure}

    \caption{Evolution of the pandemic, its drivers, and model fit, September 2020 to May 2021}
    \label{fig:pandemic_drivers_model_fit}

    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
    Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
    Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
    \ref{subsec:new_cases_fit}. For legibility reasons, all lines in
    Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
    Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that a
    value of 1 is refers to the situation at the start of our sample period and 0 means
    that all NPIs included in the index are turned on. The other lines in
    Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect of
    contact reductions, increased hygiene regulations, and seasonality. See
    Appendix~\ref{subsec:policies_seasonality} for separate plots of the three factors
    by contact type.}
\end{figure}

By this time, the set of policy instruments had become much more diverse. Around the
turn of the year, the first people were vaccinated with a focus on older age groups and
medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just
over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
started to replace regular PCR tests for staff in many medical and nursing facilities.
These had to be administered by medical doctors or in pharmacies. At-home tests approved
by authorities became available in mid-March, rapid test centers were opened and one
test per person and week was made available free of charge. Depending on the state,
customers were only allowed to enter certain stores with a recent negative rapid test
result. These developments are characteristic of many countries: The initial focus on
NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
acceptance and use of rapid tests. At broadly similar points in time, novel strains of
the virus have started to pose additional challenges.

We draw simulated samples of agents from the distribution of recorded infections in
September 2020 and use the model to predict recorded infection rates until the end of May
2021.\comment[id=J]{I do not understand} See Supplementary
Materials~\ref{sec:data_and_parameters} and \ref{sec:model} for detailed descriptions of
the data and the model, respectively. The blue line in Figure~\ref{fig:aggregated_fit}
shows our model's predictions are very close to officially recorded cases in the
aggregate. This is also true for infections by age and geographical region, which are
shown in the supplementary materials (Figures~\ref{fig:age_group_fit} and
\ref{fig:state_fit}, respectively).

The effects of various mechanisms can be disentangled due to the distinct temporal
variation in the drivers of the pandemic. Next to the stringency index, the three lines
in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions, increased hygiene
regulations, and seasonality evolved since early September for each of the three broad
contact networks. For example, a value of 0.75 for the work multiplier means that if the
environment was the same as in September (levels of infection rates, no rapid tests or
vaccinations, only the wildtype virus present), infections at the workplace would be
reduced by 25\%. The lines show the product of the effect of contact reductions,
increased hygiene regulations, and seasonality. Two aspects are particularly
interesting. First, all lines broadly follow the stringency index and they would do so
even more if we left out seasonality and school vacations (roughly the last two weeks of
October, two weeks each around Christmas and Easter, and some days in late May). Second,
the most stringent regulations are associated with the period of strong decreases in new
infections between late December 2020 and mid-February 2021. The reversal of the trend
is associated with he spread of the B.1.1.7 variant. The steep drop in recorded cases
during May 2021 is associated with at least weekly rapid tests to around 42~percent of
the population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
impacting a fall in the relative infectiousness of contacts outside of work and school.

In order to better understand the contributions  of rapid tests, vaccinations, and of
seasonality on the evolution of infections in 2021, Figure~\ref{fig:2021_scenarios_broad}
considers various scenarios. NPIs are always the same as in the baseline scenario.
Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
and three scenarios turning these factors off one by one.
Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
\citep{Shapley2016} to decompose the difference in total infections between the scenario
without any of the three factors and our main specification.\comment[id=J]{I would not
mention the word "coalition" in the notes to the figure on the Shapley values.
Combination with other channels is better. I think I would prefer to have a short section
on shapley values in the appendix (why use them, how do they work) and just link to that
in the figure note}

\begin{figure}[!tp]
    \centering

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
        \caption{{Recorded cases: 2021 scenarios}}
        \label{fig:2021_scenarios_recorded}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
        \caption{{Total cases: 2021 scenarios}}
        \label{fig:2021_scenarios_newly_infected}
    \end{subfigure}

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_area}
        % \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_channels_bar}
        \caption{Decomposition of the difference between the scenario without any of the
            three factors and the main scenario in
            Figure~\ref{fig:2021_scenarios_newly_infected}.}
        \label{fig:2021_scenarios_decomposition}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_area}
        % \includegraphics[width=\textwidth]{../figures/results/figures/full_decomposition_rapid_tests_bar}
        \caption{Decomposition of the difference between the scenario without rapid
            tests and the main scenario in
            Figure~\ref{fig:2021_scenarios_newly_infected}.}
        \label{fig:2021_scenarios_decomposition_tests}
    \end{subfigure}

    \caption{The effect of different interventions on recorded and actual infections}
    \label{fig:2021_scenarios_broad}

    \floatfoot{\noindent \textit{Note:} The blue line in
        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
        variant is introduced in the same way; vaccinations, rapid tests, and seasonality
        remain at their January levels. The other scenarios turn these three factors on
        one-by-one. The decompositions in Figures~\ref{fig:2021_scenarios_decomposition}
        and \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values
        where the individual contribution of a channel is its average contribution over
        different sizes of coalitions (combinations with other channels). The individual
        contribution to a coalition is the difference between the effect size of the
        coalition with the particular channel and without.}
\end{figure}

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active,
the final level is always the highest in case of the vaccination campaign (orange
lines). The Shapley value decomposition shows that vaccinations contribute about 15\% to
the cumulative difference between scenarios. Reasons for this are the slow start---it
took until March~24th until 10\% of the population had received their first vaccination,
the 20\% mark was reached on April 19th---and the focus on older individuals. These
groups contribute less to the spread of the disease than others due to a lower number of
contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. It is important to
note that the initial focus of the campaign was to prevent deaths and severe disease;
the case fatality was rate considerably lower during the third wave when compared to the
second (4.4\% between October and February and 1.4\% between March and June). By the end
of our study period, when first-dose vaccination rates reached around 40\% of the
population, the numbers of new cases would have started to decline.

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and recorded cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace. Nevertheless, we estimate it to be a quantitatively important factor
determining the evolution of the pandemic, explaining most of the early changes and
almost 40\% of the cumulative difference by the end of May.

The largest effect---almost one half when considering the decompositions---comes from
rapid testing. Here, it is crucial to differentiate between recorded cases and actual
cases. Additional testing means that infections are detected, which would otherwise
remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
until late April, recorded cases are higher than in the scenario where none of the three
mechanisms are turned on. Compared to the scenario with vaccinations only, this point is
reached around mid-May and it would be June for the comparison with the seasonality-only
scenario. The effect on total cases, however, is visible immediately. Despite the fact
that only a small fraction of the population performed weekly rapid tests in
March\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
week of March. 8\% in the model have ever done a rapid test at that point; 27\% of
respondents to the COSMO study reported in early March that they \textit{ever} did a
rapid test)}, new infections on April 1 would be reduced by 53\% relative to the scenario
without vaccinations, rapid tests, or seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests only (purple line in
Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the three
channels for rapid tests. Tests for pupils have the smallest effect, which is largely
explained by the relatively small number of students (Only 18\% of our population are in
the education sector (pupils, teachers etc.) versus  46\% are workers outside the
education sector) and the fact that schools did not operate at full capacity during our
period of study. Almost 40\% come from tests at the workplace. Despite the fact that
rapid tests for private reasons are phased in only late, they make up for more than half
of the total effect.\comment[id=HM]{How many tests are actually performed along each of
the three dimensions? Could this be a quantity story rather than the directed testing
story I am making up here as wishful thinking?}\comment[id=K]{See subsection
\ref{subsec:appendix_rapid_tests}.}

The reason lies in the fact that a substantial share of these tests is driven by an
elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
up on a positive test of a household member. The latter is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
Kretzschmar2020}. This is also the reason for why the number of tests performed fall at
the very end of our study period. Falling infection rates mean that there are fewer
events triggering private test demand.\comment[id=HM]{Positive rates along the three
channels might be nice here, if previous comment is not a quant
story}\comment[id=K]{See figure~\ref{fig:share_of_rapid_tests_by_reason_by_infected}}\comment[id=HM]{I think there is a story in here about
risk conditional on past exposure vs. prospective risk in gatherings, but it is too late
to work through it. We might just want to calculate the
probabilities.}\comment[id=K]{I don't understand which probabilities you mean.}

Two of the most contentious NPIs concern schools and mandates to work from home. In many
countries, schools switched to remote instruction during the first wave, so did Germany.
After the summer break, they were operating at full capacity with increased hygiene
measures, before being closed again from mid-December on. Some states started opening
them gradually in late February, but usual operation was not back until the beginning of
June. Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
school starting at Easter, at which point rapid tests had become widely available. We
estimate the realized scenario to have essentially the same effect as a situation with
closed schools. Under fully opened schools with mandatory tests, total infections would
have been 7\% higher; this number rises to 20\% without tests. These effect sizes are
broadly in line with empirical studies, e.g., \citet{Vlachos2021}\comment[id=HM]{Need a
back-of the envelope calculation} To use another metric, the effective weekly
reproduction number differs by $0.017$ and $0.045$, respectively. In light of the large
negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021} \comment[id=HM]{need to cite a couple of poor learning outcomes / mental
health papers. There must be well-published ones}\comment[id=K]{I googled and found two
that I think are at least not terrible but I also didn't find anything really well
published}---and in particular on those with low socio-economic status---these results in
conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when no test
strategy is available, trade-offs may well be different.

\begin{figure}[!tp]
    \centering

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
        \caption{{Effects of different schooling scenarios}}
        \label{fig:school_scenarios}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=0.9 \textwidth]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
        \caption{{Effects of different work scenarios}}
        \label{fig:workplace_scenarios}
    \end{subfigure}
    \vskip3ex

    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
    \label{fig:school_workplace_scenarios}

    \floatfoot{\noindent Note: Interventions start at Easter because there were no
        capacity constraints for rapid tests afterwards.}

\end{figure}

Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of employees receiving tests, testing at the workplace has larger effects than changing requirements to work from home. Whether the share of workers working at the usual workplace is reduced or increased by ten percent hardly has an effect on infection rates. Making testing mandatory twice a week---assuming independet compliance by employers and workers or 95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by employers to the level of March would have increased them by the same amount.

Despite these large effects, the results on testing likely understate the benefits.
Disadvantaged groups of the population are less likely to be reached by testing
campaigns;\comment[id=HM]{seems obvious, but I did not find anything. COSMO data? We
should get the raw data, btw.} at the same time they have a higher risk to contract
CoViD-19.\citet{KochInstitut2021a} Mandatory tests at school and at the workplace will
extend more into these groups. The same goes for individuals who exhibit a low level of
compliance with CoViD-19-related regulations.

Our analysis\comment[id=J]{The conclusion should be slightly longer and more separated from the rest} has shown that in the transition path to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A day of the winter lockdown in Germany is estimated to have
cost 6.25~billion~Euros \citep{Dorn2020b}.\comment[id=K]{This is a prediction by the ifo
Instute back from March 2020 assuming a 2 month lockdown. I did not find anything citable
that is more recent. I did find
\href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A
quote from ifo's Timo Wollmershäuser: „Wir liegen zwischen zwei bis zweieinhalb
Milliarden Euro pro Woche“} for the winter lockdown. That would be much less than
calculated in March.} In early June 2021, retail prices for self-administered rapid tests
are frequently less than 1.5~Euro.\footnote{Cheapest on \href{https://amazon.de}{Amazon}
on June 11th, 2021.}