### 2021-06-16 paper/model/estimated_parameters.tex

\subsection{Estimated Parameters}
\label{subsec:estimated_params}

%%% The free parameters are the infection probabilities by contact category, one hygiene
%%% multiplier for school and work contacts after November 2nd 2020, approx. ten parameters
%%% that govern the reduction of other contacts\comment[id=K]{Must be updated when the
%%% estimation is finished.}, the number of extra contacts during holidays, the share of
%%% detected cases around holidays and the fade in speed of the rapid tests as well as one
%%% parameter that governs the import of B.1.1.7 cases in January 2021.

\FloatBarrier

% infection probabilities

To calibrate infection probabilities outside of the model, it would be important to know
the exact duration and distance of each contact type as well as viral loads. Since this
is not available in any data set, we estimate those parameters inside the model with the
method of simulated moments \citep{McFadden1989} by minimizing the distance between
simulated and observed infection rates. Since our model includes a lot of randomness, we
average simulated infection rates over several model runs.

We use data for Germany from October 2020 until June 2021. We do not use
earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
limited and therefore it would be difficult to find good initial conditions for our
simulations. In addition during the summer the case numbers were extremely low. This
could lead to the epidemic going extinct in our simulation. Additionally, our model does
not include international travel or other imports of cases. These would be important but
difficult to model during the summer months.

To avoid over-fitting and simplify the numerical optimization problem, we only allow for
five different probabilities: 1) for contacts in schools 2) contacts in preschools and
nurseries. 3) for work contacts. 4) for households. 5) for other contacts.

The infection probabilities estimated from our model are as follows: For household
contacts (each individual meets all his or her household members every day) the infection
probability is 10\%, for school contacts (all students and teachers meet all other
students and teachers that attend school that day and there are three classes taking
place each day) the probability of infection is 1.2\%, for preschools and nurseries that
infection probability is 0.5\%. Work contacts carry a risk of 14.5\% and other activities
(including leisure activities such as meeting friends) carry the highest risk with
15.875\%.\comment[id=HM]{Are these numbers current? In any case, a table would substitute for this and the previous paragraph. Need to mention B.1.1.7 alongside.}\comment[id=K]{Missing to add data for the other estimated parameters (import of b117 etc)}