### 2021-06-17 paper/model/estimated_parameters.tex
Put estimated infection probabilities into a table.

\subsection{Estimated Parameters}
\label{subsec:estimated_params}

%%% The free parameters are the infection probabilities by contact category, one hygiene
%%% multiplier for school and work contacts after November 2nd 2020, approx. ten parameters
%%% that govern the reduction of other contacts\comment[id=K]{Must be updated when the
%%% estimation is finished.}, the number of extra contacts during holidays, the share of
%%% detected cases around holidays and the fade in speed of the rapid tests as well as one
%%% parameter that governs the import of B.1.1.7 cases in January 2021.

\FloatBarrier

We estimate parameters that could not be calibrated inside the model with the
method of simulated moments \citep{McFadden1989} by minimizing the distance between
simulated and observed infection rates. Since our model includes a lot of randomness, we
average simulated infection rates over several model runs.

We use data for Germany from October 2020 until June 2021. We do not use
earlier periods for three reasons. Firstly, in the beginning PCR tests were highly
limited and therefore it would be difficult to find good initial conditions for our
simulations. In addition during the summer the case numbers were extremely low. This
could lead to the epidemic going extinct in our simulation. Additionally, our model does
not include international travel or other imports of cases. These would be important but
difficult to model during the summer months.

% infection probabilities

To avoid over-fitting and simplify the numerical optimization problem, we only allow for
five different probabilities: 1) for contacts in schools 2) contacts in preschools and
nurseries. 3) for work contacts. 4) for households. 5) for other contacts.

The infection probabilities estimated from our model are as shown in
Table~\ref{tab:infection_probs}.\comment[id=K]{Missing to add data for the other
estimated parameters (import of b117 etc)}\comment[id=K]{Update table after the new
estimation is finished}

\begin{table}[tb]
    \caption{Estimated Infection Probabilities}
    \label{tab:infection_probs}
    \centering
    \begin{tabular}{ll}
        \toprule
        Contact Type & Infection Probability \\
        \midrule
        preschool and nursery & 00.500\% \\
        school & 01.200\% \\
        household & 10.000\% \\
        work & 14.500\% \\
        other & 15.875\% \\
        \bottomrule
    \end{tabular}
    \tablenotes{School contacts take place Monday through Friday. All students and
teachers meet all other students and teachers that attend school that day and there are
three classes taking place each day. Work contacts can be daily, weekly or non recurrent.
Other contacts can be daily, weekly or non recurrent. We do not differ the infection
probability depending on how often a contact takes place.}
\end{table}