### 2021-06-21 src/paper.tex
More adjustments for arxiv.

\pdfoutput=1

\input{paper/preamble}

\begin{document}

\selectlanguage{USenglish}
\frenchspacing

\input{paper/titlepage}

\maketitle
\pagenumbering{arabic}
\setcounter{page}{1}

\newpage

\input{paper/report}

\phantomsection % Needed for the bibliography bookmark to point to the correct page

\begin{refcontext}[sorting=nyt]  % Sort BIBLIOGRAPHY by alphabet (while CITATIONS are sorted by year)
\printbibliography[heading=bibintoc]
\end{refcontext}

\clearpage

\begin{center}
    {
        \Large 
        Supplementary Material for\\[2ex]

        \textbf{\papertitle}
    }
    \vskip4ex
    {
        \large
        Janoś Gabler,

        Tobias Raabe,

        Klara Röhrl,

        Hans-Martin von Gaudecker

    }
\end{center}

\clearpage

\begin{appendices}
    \begin{refsection}

    \input{paper/appendix_data}

    \FloatBarrier

    \input{paper/appendix_model}

    \input{paper/appendix_results}

    \printbibliography[heading=bibintoc]

    \end{refsection}
\end{appendices}

\end{document}