### 2021-06-22 src/paper/abstract.tex

\noindent
In order to slow the spread of the CoViD-19 pandemic, governments around the world have
enacted a wide set of policies limiting the transmission of the disease. Initially,
these focused on non-pharmaceutical interventions; more recently, vaccinations and
large-scale rapid testing have started to play a major role. The objective of this study
is to explain the quantitative effects of these policies on determining the course of
the pandemic, allowing for factors like seasonality or virus strains with different
transmission profiles. To do so, the study develops an agent-based simulation model,
which is estimated using data for the second and the third wave of the CoViD-19 pandemic
in Germany. The paper finds that during a period where vaccination rates rose from 5\%
to 40\%, rapid testing had the largest effect on reducing infection numbers. Frequent
large-scale rapid testing should remain part of strategies to contain CoViD-19; it can
substitute for many non-pharmaceutical interventions that come at a much larger cost to
individuals, society, and the economy.

\vspace{1cm}
\noindent \textbf{JEL Classification:} C63, I18

\noindent \textbf{Keywords:} CoViD-19, agent based simulation model, rapid testing,
non-pharmaceutical interventions