### 2021-06-28 src/paper/model/param_tables.tex
Very first version of adding parameter tables.

\subsection{Overview Model Parameters} % (fold)
\label{sub:param_tables}

\begin{table}
\centering
\caption{contacts_and_matching}
\label{tab:contacts_and_matching}
\input{../bld/param_tables/contacts_and_matching}
\end{table}

\begin{table}
\centering
\caption{policies}
\label{tab:policies}
\input{../bld/param_tables/policies}
\end{table}

\begin{table}
\centering
\caption{infection_probs}
\label{tab:infection_probs}
\input{../bld/param_tables/infection_probs}
\end{table}

\begin{table}
\centering
\caption{disease_model}
\label{tab:disease_model}
\input{../bld/param_tables/disease_model}
\end{table}

\begin{table}
\centering
\caption{pcr_testing_and_case_detection}
\label{tab:pcr_testing_and_case_detection}
\input{../bld/param_tables/pcr_testing_and_case_detection}
\end{table}

\begin{table}
\centering
\caption{rapid_testing}
\label{tab:rapid_testing}
\input{../bld/param_tables/rapid_testing}
\end{table}

\begin{table}
\centering
\caption{behavioral_response}
\label{tab:behavioral_response}
\input{../bld/param_tables/behavioral_response}
\end{table}

\begin{table}
\centering
\caption{vaccinations}
\label{tab:vaccinations}
\input{../bld/param_tables/vaccinations}
\end{table}