### 2021-06-30 src/paper/model/param_tables.tex
Merge tables.

\clearpage

\begin{landscape}

\subsection{Overview Model Parameters} % (fold)
\label{sub:param_tables}

\FloatBarrier

\begin{table}[htb]
\centering
\caption{Contacts, Matching and Policies}
\label{tab:contacts_and_matching}
\input{../bld/param_tables/contacts_matching_policies}
\end{table}

\begin{table}[htb]
\centering
\caption{Infection Probability}
\label{tab:infection_probs}
\input{../bld/param_tables/infection_probs}
\end{table}

\begin{table}[htb]
\centering
\caption{Disease and Vaccination Model}
\label{tab:disease_and_vacc_model}
\input{../bld/param_tables/disease_model}
\end{table}

\begin{table}[htb]
\centering
\caption{Rapid Testing}
\label{tab:rapid_testing}
\input{../bld/param_tables/rapid_testing}
\end{table}

\begin{table}[htb]
\centering
\caption{PCR Testing, Case Detection and Behavioral Response}
\label{tab:pcr_testing_and_case_detection}
\input{../bld/param_tables/pcr_testing_and_case_detection}
\end{table}

\FloatBarrier

\end{landscape}

\FloatBarrier