### 2021-07-09 src/paper/appendix_results.tex
Refactor the results section. Add random rapid test scenarios and robustness check scenarios (no descriptions yet).

\section{Additional Results}\comment[id=K]{Philipp suggests to do a sensitivity analysis
for the seasonality}
\label{sec:additional_results}

\input{paper/results/fit}

\input{paper/results/share_known_cases}

\input{paper/results/rapid_tests_statistics}

\input{paper/results/scenarios}