### 2021-07-10 src/paper/appendix_data.tex
Move data text to different text files. Arange number of contacts figure.

\section{Data and Parameters}
\label{sec:data_and_parameters}

The model is described by a large number of parameters that govern the number of
contacts a person has, the likelihood of becoming infected on each contact, the
likelihood of developing light or strong symptoms or even dying from the disease as well
as the duration each stage of the disease takes.

Most of these parameters can be calibrated from existing datasets or the medical
literature or calibrated from surveys and empirical datasets.

\input{paper/data/medical_parameters}

\input{paper/data/synthetic_population}

\input{paper/data/number_of_contacts}

\input{paper/data/assortativity}

\input{paper/data/policies}