### 2021-07-10 src/paper/data/rapid_test_demand.tex
Continue extending the fit section.

\subsection{Rapid Test Demand}
\label{subsec:rapid_test_demand}

In our model, there are five reasons why rapid tests are done:\comment[id=J]{ Add a
    section on how we calibrate rapid test demand; Mainly describe the datapoints we have
    and say that we usually interpolate linearly in between data points. (Only exception
    to that is private rapid test demand, which we fit to data)
}

\begin{enumerate}
    \item someone plans to have work contacts
    \item someone is an employee of an educational facility or a school pupil
    \item a household member has tested positive or developed symptoms
    \item someone has developed symptoms but has not received a PCR test
    \item someone plans to participate in a weekly non-work meeting
\end{enumerate}

%%% \subsubsection{work rapid tests}

For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April)
that 60\% of workers who receive a test offer by their employer regularly use it. We
assume this share to be time constant.

In addition, there are some surveys that allow us to trace the expansion of employers who
offer tests to their employees. Mid march, 20\% of employers offered tests to their
employees \citep{DIHK2021}. In the second half of March, 23\% of employees reported being
offered weekly rapid tests by their employer \citep{Ahlers2021}. This share increased to
61\% until the first days of April \citep{BMWI2021, IZA2021}.

Until mid April 72\% of workers were expected to receive a weekly test offer
\citep{BMWI2021, IZA2021}. However, according to surveys conducted in mid April
\citep{Betsch2021}, less than two thirds of individuals with work contacts receive a test
offer. Starting on April 19th employers were required by law to provide two weekly tests
to their employees \citep{Bundesanzeiger2021}. We assume that compliance is incomplete
and only 80\% of employers actually offer tests.

%%\subsubsection{educ rapid tests}

We assume that employees in educational facilities start getting tested in 2021 and that
by March 1st 30\% of them are tested weekly. The share increases to 90\% for the week
before Easter. At that time both Bavaria \citep{STMGP2021} and Baden-Württemberg
\citep{MinisteriumKultus2021} were offering tests to teachers and North-Rhine Westphalia
\citep{SchulministeriumNRW2021} and Lower Saxony \citep{NSMK2021} were already testing
students and tests for students and teachers were already mandatory in Saxony
\citep{SMK2021}. After Easter we assume that 95\% of teachers get tested twice per week.

Tests for students started later \citep{MinisteriumKultus2021, SchulministeriumNRW2021}
so we assume that they only start in February and only 10\% of students get tested by
March 1st. Relying on the same sources as above we approximate that by the week before
Easter this share had increased to 40\% \citep{SchulministeriumNRW2021}.

After Easter the share of students receiving twice weekly tests is set to 75\%. This is
based on tests becoming mandatory in Bavaria \citep{BayerischeStaatskanzlei2021} and
North Rhine-Westphalia \citep{SchulministeriumNRW2021b} after their Easter breaks and
on the 19th in Baden-Württemberg \citep{KMBaWue2021}.

%%\subsubsection{private rapid tests}

To limit our degrees of freedom, we only have one parameter that governs how many
individuals do a rapid test because of any of the private demand reasons (own symptoms
but no PCR test, planned weekly leisure meeting or a symptomatic or positively tested
household member).

We assume that there is no private rapid test demand until March when both the citizens'
tests and rapid tests for lay people started to become available
\citep{Bundesanzeiger2021a, Bundesregierung2021} and other access to rapid tests was very
limited.

According to the COSMO study \citep{Betsch2021a} 63\% would have been willing to take a
test in the round of 23rd of February 2021 when an acquaintance would have tested
positive. Since this is only asking for willingness not actual behavior and the demand
when meeting with friends is very likely lower, we take this as the upper bound of
private rapid test demand which is reached in the beginning of May. To cover that many
people are likely to have sought and done their first rapid test before the Easter
holidays to meet friends or family, we let the share of individuals doing rapid tests in
that time increase more rapidly than before and after. By end of March 25\% of
individuals would do a rapid test due to a private reason.

All shares of individuals who would take a rapid test if the conditions were met can be
seen in Figure~\ref{fig:rapid_test_demand}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand_shares}
    \caption{\textbf{Share of Individuals Doing a Rapid Test.}}
    \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
    planning to have education contacts, work contacts, developing symptoms without
    access to a PCR test, having a household member with a positive test or symptoms. In
    each case whether a rapid test is done depends on how long it has been since the
    individual's last rapid test and her individual compliance parameters. As an example,
    take a worker in May. In that time workers are encouraged to test themselves twice
    weekly but there is no general requirement to test themselves. If the worker has not
    done a test within the last four days in our model she will demand a test if her
    (time-constant) compliance parameter belongs to the upper 60\% in the population.}
    \label{fig:rapid_test_demand}
\end{figure}

\begin{figure}[ht]
    \centering
    \caption{Share of Individuals With Rapid Tests}
    \label{fig:share_ever_rapid_test}
    \begin{subfigure}{.55\textwidth}
        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
        \caption{Share of Individuals That Have Ever Done a Rapid Test}
    \end{subfigure}%
    \begin{subfigure}{.55\textwidth}
        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
        \caption{Share of Individuals Having Done a Rapid Test in the Last Week}
    \end{subfigure}
    \label{fig:share_rapid_test_last_week}
    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
        have ever done a rapid test or done a rapid test within the last week in our
        simulations to the shares reported in the
        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
        have ever done a rapid test. The right panel compares the share of individuals
        who have done a rapid test within the last seven days in our simulation compared
        to the share reporting to have done at least weekly rapid tests in the last four
        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
        conservative. The overall share is below that in the study. We fit the share of
        weekly tests quite exactly. However, the study only covers adults while our share
        also includes children who are tested very regularly when attending school.}
\end{figure}

 \comment[id=K]{J: Explain that we don't fit the share ever tested that well because our rapid test compliance is completely fixed. }

\FloatBarrier