### 2021-07-11 src/paper/data/medical_parameters.tex

\subsection{Medical Parameters}
\label{sec:medical_params}

This section discusses the medical parameters used in the model, their sources and how we
arrived at the distributions used in the model.\footnote{Additional information can be
found in the
\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
documentation}.} See Figure~\ref{fig:model_disease_progression} for a summary of our
disease progression model.

% Duration until Infectiousness and Symptoms ----------------------------------

% Latency Period
The first medical parameter we need is the length of the period between infection and the
start of infectiousness, the so called latent period. \cite{Zhao2021} estimate the latent
period to last 3.3 days (95\% CI: 0.2, 7.9). In line with this estimate our latent period
lasts one to five days with a slightly lower average of 2 days.

% Asymptomatic Share
Once individuals become infectious a share of them goes on to develop symptoms while
others remain asymptomatic. We rely on data by \cite{Davies2020} for the age-dependent
probability to develop symptoms. It varies from 25\% for children and young adults to
nearly 70\% for the elderly.

% Incubation Period
The incubation period is usually estimated to be two to twelve days. A meta analysis by
\citet{McAloon2020} comes to the conclusion that ``The incubation period distribution may
be modeled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\%
CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For
simplicity we discretize this distribution into four bins.

% Length of the presymptomatic stage
Similar to \citet{Peak2020} and in line with \citet{He2020} we set the length of the
presymptomatic stage to be one to two days.

% Duration of Infectiousness ----------------------------------------------------------

We assume that the duration of infectiousness is the same for both symptomatic and
asymptomatic individuals as evidence suggests little differences in the transmission
rates between symptomatic and asymptomatic patients (\citet{Yin2020}) and that the viral
load between symptomatic and asymptomatic individuals are similar (\citet{Zou2020},
\citet{Byrne2020}, \citet{Singanayagam2020}). Our distribution of the duration of
infectiousness is based on \citet{Byrne2020}. For symptomatic cases they arrive at zero
to five days before symptom onset (see their figure 2) and three to eight days of
infectiousness afterwards.\footnote{Viral loads may be detected much later but eight days
seems to be the time after which most people are culture negative, as also reported by
\citet{Singanayagam2020}.} Thus, we arrive at zero to 13 days as the range for
infectiousness among individuals who become symptomatic (see also figure 5).

% Duration of Symptoms ----------------------------------------------------------

We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3,
Panel~2]{Bi2020} for the duration of symptoms for non-ICU requiring symptomatic cases. We
only disaggregate by age how likely individuals are to require intensive
care.\footnote{The length of symptoms is not very important in our model given that
individuals mostly stop being infectious before their symptoms cease.}

% Time from Symptom Onset to Admission to ICU ------------------------------------------

For the time from symptom onset until need for intensive care we rely on data by
\cite{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.

For those who will require intensive care we follow \citet{Chen2020} who estimate the
time from symptom onset to ICU admission as $8.5 \pm 4$ days. This aligns well with
numbers reported for the time from first symptoms to hospitalization:
\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of four. We assume
that the time between symptom onset and ICU takes four, six, eight or ten days with equal
probabilities.

% Death and Recovery from ICU ----------------------------------------------------------

We take the survival probabilities and time to death and time until recovery from
intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}. They report
time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. To
match this approximately we discretize that 41\% of individuals who will die from
Covid-19 do so after one day in intensive care, 22\% day after twelve days, 29\% after 20
days and 7\% after 32 days. Again, we rescale this for every age group among those that
will not survive. For survivors the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}
reports a mean duration of 18.8 days until recovery and a standard deviation of 12.21
days. We discretize this such that of those who recover in intensive care, 22\% do so
after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

\FloatBarrier