### 2021-07-11 src/paper/appendix_data.tex

\section{Data and Parameters}
\label{sec:data_and_parameters}

The model is described by a large number of parameters that govern the number of contacts
a person has, the reduction in contacts due to NPIs, the demand for rapid tests and PCR
tests, the likelihood of becoming infected on each contact, the likelihood of developing
light or strong symptoms or even dying from the disease as well as the duration each
stage of the disease takes.

Many of these parameters can be calibrated from the medical literature or calibrated from
surveys and other empirical datasets.

\input{paper/data/medical_parameters}

\input{paper/data/synthetic_population}

\input{paper/data/number_of_contacts}

\input{paper/data/assortativity}

\input{paper/data/policies}

\input{paper/data/rapid_test_demand}

\input{paper/data/share_known_cases}