### 2021-07-11 src/paper/data/policies.tex
Make policy data section ready for review.

\subsection{Policies}
\label{subsec:policies}

Our policies usually affect one of our four contact types: households, education, work
and other contacts.
% households
For households we assume that the individuals' contacts in their households do not change
over our estimation period.
% educ models
For nurseries, preschools and schools we implement vacations as announced by the German
federal states as well as school closures, emergency care and rotating schooling
schedules where only one half of students attends every other week or day.% For the moment we ignore that lack of childcare leads working parents to stay home.
An approximation of the share of contacts still taking place with the different school
regulations can be found in Figure~\ref{fig:school_multiplier}.
% work
For work contacts we use the reductions in work mobility reported by the Google Mobility
Data \citep{Google2021} to calibrate our work policies. Reductions in work contacts are
not random but governed through a work contact priority where the policy changes the
threshold below which workers stay home. Figure \ref{fig:work_multiplier} shows the share
of workers that go to work over time at the federal German level. We use the data on the
state level to account for local holidays and differences in state regulations.
% hygiene umltiplier
In addition, for both work and school contacts we assume that hygiene measures (such as
masks, ventilation and hand washing) became more strict and more conscientiously observed
in November, leading to a reduction of 33\% in the number of contacts with the potential
to transmit Covid-19.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/school_multiplier_comparison}
        \caption{{School Attendance}}
        \label{fig:school_multiplier}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/work_multiplier_since_sep}
        \caption{Work Attendance}
        \label{fig:work_multiplier}
    \end{subfigure}

    \vskip3ex

    \caption{The Contact Reduction Effects of School and Work Attendance Policies}

    \floatfoot{\noindent \textit{Note:} The left figure shows the approximate share of
    school contacts taking place with and without vacations factored in. The dates on
    which schools have vacation are decided at the state level. Vacations are directly
    implemented in our model with no school contacts taking place on weekends and during
    vacations (by federal state) just like the schooling mode (full operation, emergency
    care, rotating schemes with half class sizes etc.). The figure is thus only an
    illustration that shows the approximate share of contacts taking place compared to
    the pre-pandemic level with and without vacations. The difference between the lines
    show when vacations take place and to what degree. For example all states have fall
    vacations but the timing varies strongly between states while all states close
    schools during the Christmas holidays. The right figure shows the work mobility as
    reported by \cite{Google2021}. We take this as a proxy of the share of workers who
    still have physical work contacts. The figure interpolates over weekends as we handle
    weekend effects through information on work on weekends in the German census data we
    use. The figure shows the share for Germany as a whole. To capture the effect that
    local policies, school vacations, etc. have on work contacts we use the data on the
    state level to determine which workers go to work depending on the state they live
    in.}
    \label{fig:multipliers}
\end{figure}

Lastly, for the other contacts category we could not calibrate the policies from data but
estimated the policy effects. The estimation and values are detailed in
Section~\ref{subsec:estimated_params} and Figure~\ref{fig:other_multiplier}.

\FloatBarrier