### 2021-07-11 src/paper/data/share_known_cases.tex
Make data section on share known cases ready for review. Add figures showing the parameter demand shares in different robustness scenarios.

\subsection{Share of Detected Cases}
\label{subsec:data_share_known_cases}

 \comment[id=K]{Stoye called this the ascertainment rate if I remember right. Maybe we should replace this everywhere}

One important feature of our model is that we distinguish between undetected and detected
cases and that we model which cases are detected and which are not (see
Section~\ref{sub:testing} for a detailed description for how we model both rapid and PCR
tests). For our model it is important to have an estimate for the share of cases that is
detected in the absence of rapid tests. For this we rely on the Dunkelzifferradar Project
\citep{Dunkelzifferradar2020} which relies on estimates of the case fatality rate to
estimate the number of total cases given the number of CoViD-19 deaths which are assumed
to be perfectly observable. For 2020, we follow the reported share of detected cases
quite closely only interpolating the phase of November 2020 where the testing policy was
changed to give more priority to vulnerable groups. This lead to more tests being done
among older age groups which are also at the highest risk of dying from CoViD-19 and an
overestimation of the share of detected cases.
% After Christmas 2020
Since vaccinations started after Christmas 2020 and these were predominantly given to
nursing homes in the beginning and other vulnerable groups in spring, we expect the
relationship between deaths and the number of total infections to change rapidly in 2021.
This is why we stop using the share of detected cases estimated by the Dunkelzifferradar
after Christmas. Instead, we assume that the share of detected cases would have
stayed the same in the absence of rapid tests. Thus, we also achieve in our model an
increase in the share of detected cases but this is driven from inside our model through
increased rapid testing and follow-up on positive rapid tests (see
Section~\ref{subsec:results_share_known_cases} and \ref{sub:testing}).
% Christmas and Easter
Lastly, we model reductions in the share known cases due to the two major holidays in our
simulation period, Christmas and Easter. During both holidays many laboratories did not
process tests and most physicians' offices were closed, leading to less PCR tests and
short and large drops in the share of known cases. The resulting share of detected cases
in the absence of rapid tests is shown in Figure~\ref{fig:share_known_cases_data}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/assumed_overall_share_known_cases}
  \caption{Share of Detected Cases in the Absence of Rapid Tests}
  \label{fig:share_known_cases_data}
  \floatfoot{\noindent \textit{Note:} The figure shows the share of cases that is
  reported as an official case via PCR confirmation. We use the overall share of known
  cases that was estimated through the case fatality ratio by the
  \citet{Dunkelzifferradar2020} for all of 2020 and
  then assume it to be constant as vaccinations of the elderly strongly affect the case
  fatality rate which the project does not account for. Starting in 2021 in addition to
  the overall numbers of detected cases through symptoms and the share known cases, cases
  are also detected through confirmation of positive rapid tests which happens
  endogenously inside the model. For the public holidays of Christmas and Easter we lower
  the share of detected cases as fewer PCR tests are available during public holidays.}
\end{figure}

\FloatBarrier