### 2021-07-12 src/paper/report.tex
Add notation from the parameter tables to the report main text (mostly as comments).

Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many
months---and often years---away from reaching vaccination levels that would end the
pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
of utmost importance to employ an effective mix of strategies for containing the virus.
The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
to reduce contacts between individuals. While this has allowed some countries to sustain
equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
theoretical equilibrium at low case numbers which is sustained with
test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
over time. Containment measures have become increasingly diverse and now include rapid
testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
influence of seasonal patterns or of more infectious virus strains are well understood
in quantitative terms.

This paper develops a quantitative model incorporating these factors simultaneously. The
framework allows to combine a wide variety of data and mechanisms in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany, where new infections fell by almost 80\% during the May 2021. Our analysis shows
that, aside from seasonality, frequent and large-scale rapid testing caused the bulk of
this decrease, which is in line with prior predictions \citep{Mina2021}. We conclude that
it should have a large role for at least as long as vaccinations have not been offered to
an entire population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The model
most closely related to ours is described in \citet{Hinch2020}.} Each contact between an
infectious individual and somebody susceptible to the disease bears the risk of
transmitting the virus\comment[id=K]{infection probability per contact for contact type
$c$: $\beta_{c}$}. Contacts occur in up to four networks: Within the household, at work,
at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts\comment[id=K]{probability to have $n$ contacts
of contact type $c$: $\eta_{c,\:n}$} from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
depend on age and geographical location (see Supplementary Material
Section~\ref{subsec:assortativity}).\comment[id=K]{Share of contacts of age group and
county of i to meet someone of j's age group and county: $\alpha_{work,\:age_i,
age_j,\:county_i,\:county_j}$}

The four contact networks are chosen so that the most common NPIs can be modeled in great
detail. NPIs affect the number of contacts or the risk of transmitting the disease upon
having physical contact. The effect of different NPIs will generally vary across contact
types. For example, a mandate to work from\comment[id=K]{$\rho_{attend,\,t}$} home will
reduce the number of work contacts to zero for a fraction of the working population.
Schools and daycare can be closed entirely, operate at reduced capacity---including an
alternating schedule---, or implement mitigation measures like masking requirements or
air filters \citep{Lessler2021}. Curfews\comment[id=K]{would enter into $\rho_{other\:,t}$} may reduce the number of contacts in
non-work/non-school/non-work settings. In any setting, measures like masking requirements
would reduce the probability of infection associated with a contact \citep{Cheng2021}.\comment[id=K]{$\rho_{hygiene}$}

\begin{figure}   % Figure 1
    \centering

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/model-graph-top-left}
        \caption{{Model for contacts and infections}}
        \label{fig:model_contacts_infections}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/model-graph-top-right}
        \vskip5ex

        \caption{Disease progression}
        \label{fig:model_disease_progression}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering

        \includegraphics[width=\textwidth]{figures/model-graph-bottom-left}
        \caption{{Model for rapid tests}}
        \label{fig:model_rapid_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/model-graph-bottom-right}
        \vskip5ex

        \caption{{Translating all infections to recorded ones}}
        \label{fig:model_official_cases}
    \end{subfigure}

    \caption{Model description}
    \label{fig:model_description}

    \floatfoot{\noindent \textit{Note:}
        A~description of the model can be found in Supplementary
        Material~\ref{sec:model}. Figure~\ref{fig:model_contacts_infections} shows the
        influence on an agent's contacts to other agents. Demographic characteristics set
        the baseline number of contacts in different networks ($\eta$). She may reduce
        the number of contacts due to NPIs, showing symptoms, or testing positively for
        SARS-CoV-2 ($\tau$). Infections may occur when a susceptible agent meets an
        infectious agent; the probability depends on the type of contact ($\beta_c$), on
        seasonality ($\kappa_{c}$), and on NPIs ($\rho_{c,\:t}$). If infected, the infection
        progresses as depicted in Figure~\ref{fig:model_disease_progression}. If rapid
        tests are available, agents' demand is modeled as in
        Figure~\ref{fig:model_rapid_tests}. All reasons trigger a test only for a
        fraction of individuals depending on an individual compliance parameter; the
        thresholds for triggering test demand differ across reasons and they may depend
        on calendar time ($\pi_{c,\:t}$ and $\tau_{c,\:t}$).
        Figure~\ref{fig:model_official_cases} shows the model of translating all
        infections in the simulated data to age-specific recorded infections. The model
        uses data on the aggregate share of recorded cases ($\psi$), on the share of
        positive PCR tests triggered by symptoms ($\chi_{symptom}$), and on the false
        positive rate of rapid tests ($p_{positive|infected,\;i,\:t}$). The lower part of
        the graph is relevant only for periods where rapid tests are available.}
    \end{figure}

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
age\comment[id=K]{$\zeta_{a}$}. A possible infection progresses as shown in
Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period of
infection without being infectious or showing symptoms, being infectious (presymptomatic
or asymptomatic), showing symptoms, requiring intensive care, and recovery or death
\citep[similar to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The
probabilities of transitioning between these states depend on age;\comment[id=K]{$\gamma$
for all transition probabilities} their duration is random within intervals calibrated to
the medical literature (for a detailed description see Section~\ref{sec:medical_params}).
Conditional on the type of contact, infectiousness is independent of age
\citep{Jones2021}.\comment[id=K]{would be $\beta_{c}$}

The model includes several other features, which are crucial to describe the evolution of
the pandemic in 2020-2021. New virus strains with different profiles regarding
infectiousness\comment[id=K]{$\sigma_v$} can be introduced. Agents may receive a
vaccination. With a probability of 75\% \citep{Hunter2021}, vaccinated agents become
immune and they do not transmit the virus \citep{Petter2021, LevineTiefenbrun2021,
Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the second
dose of the Biontech/Pfizer vaccine, which is most commonly used in Germany. We choose it
because our model neither includes booster shots, nor does it allow vaccinated
individuals who became immune to transmit the disease\citep{Petter2021,
LevineTiefenbrun2021, Pritchard2021}. If anything, these assumptions would overstate the
effect of vaccines for our study period. This would be different if a large fraction of
vaccinated individuals had received a second dose already.} During the vaccine roll-out,
priority may depend on age and occupation.\comment[id=K]{Maybe mention that we have
refusers?}

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
some days to be processed\comment[id=K]{$\gamma_{PCR,\:d}$ where $d=$ number of days} and
there are aggregate capacity constraints throughout. In contrast, rapid antigen tests
yield immediate results; after a phase-in period, all tests that are demanded will be
performed. Specificity\comment[id=K]{$p_{negative|not\:infected}$} and
sensitivity\comment[id=K]{"$p_{positive|infected,\;i,\:t}$", where $i$ is the individual
and $t$ is the date to determine time until/since infectiousness} of these tests is set
according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
timing of the test relative to the onset of infectiousness.
Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
require staff and students to be tested regularly.\comment[id=K]{"$\pi_{teacher,\:t}$"
etc. } Rapid tests may be offered by employers to on-site workers. Individuals may demand
tests for private reasons, which include having plans to meet other people, showing
symptoms of CoViD-19, and because a household member tested positively for the virus. We
endow each agent with an individual compliance parameter. This parameter determines
whether she takes up rapid tests.\footnote{A positive test result (of either kind), as
well as symptoms leads most individuals to reduce their contacts; this is why tests
impact the actual contacts in Figure~\ref{fig:model_description}.}

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group\comment[id=K]{Should we
add all our sources here?}. Mobility data is used to model the evolution of work
contacts. School and daycare policies can be incorporated directly from official
directives. Administrative records on the number of tests, vaccinations by age and
region, and the prevalence of virus strains are generally available. Surveys may ask
about test offers, propensities to take them up, and past tests. Other studies' estimates
of the seasonality of infections can be incorporated directly. The remaining
parameters---most notably, these include infection probabilities by contact network and
the effects of some NPIs, see Supplementary Material~\ref{subsec:estimated_params}---will
be chosen numerically so that the model matches features of the data \citep[see][for the
general method, also described in Supplementary Material
\ref{subsec:estimated_param}]{McFadden1989}. In our application, we keep the number of
free parameters low in order to avoid overfitting. The data features to be matched
include official case numbers for each age group and region, deaths, and the share of the
B.1.1.7 strain.

The main issue with official case numbers is that they will contain only a fraction of
all infections. We thus model official cases as depicted in
Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
detected cases\comment[id=K]{$\psi$} and use data on the share of PCR tests administered
because of CoViD-19 symptoms\comment[id=K]{$\chi_{symptom}$}. As the share of
asymptomatic individuals varies by age group, this gives us age-specific shares (see
Figure~\ref{fig:share_known_cases_by_age_group} for the share of detected cases by age
group over time in our model). Our estimates suggest that---in the absence of rapid
testing---the detection rate is 80\% higher on average for individuals above age 80
compared to school age children. Once rapid test become available, confirmation of a
positive result\comment[id=K]{prob of demanding confirmation is $\chi_{confirmation}$} is
another reason leading to positive PCR tests.

The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
never use it directly. For legibility reasons, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value one represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of ten. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November before rising again in December, prompting the most stringent lockdown to this
date. Schools and daycare centers were closed, so were customer-facing businesses except
for grocery and drug stores. From the peak of the second wave just before Christmas until
the trough in mid-February, newly detected cases decreased by almost three quarters. The
third wave in the spring of 2021 is associated with the B.1.1.7 strain, which became
dominant in March (Figure~\ref{fig:share_b117}) and stayed dominant for our entire
simulation period.\footnote{B.1.617.2 (Delta) reached Germany in April but still
accounted for less than 5\% of cases at the end of our simulation period.}
In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement
stores were allowed to open again to the public. There were many changes in details of
regulations afterwards, but they did not change the overall stringency index.

\begin{figure}[!tp]   % Figure 2
    \centering

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_new_known_case}
        \caption{{Recorded cases: Empirical and simulated}}
        \label{fig:aggregated_fit}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/stringency2_with_seasonality}

        \caption{{Stringency of NPIs and infectious contacts}}
        \label{fig:stringency_infectious_contacts}
    \end{subfigure}

    \vskip3ex

    \begin{subfigure}[b]{0.475\textwidth}
        \centering

        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_b117}

        \caption{Fraction of B.1.1.7 strain}
        \label{fig:share_b117}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering

        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week_and_vaccinated}

        \caption{{Tests and vaccinations}}
        \label{fig:antigen_tests_vaccinations}
    \end{subfigure}

    \caption{Evolution of the pandemic, its drivers, and model fit, September 2020 to May 2021}
    \label{fig:pandemic_drivers_model_fit}

    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
        Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
        \ref{subsec:fit_results}. For legibility reasons, all lines in
        Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that a
        value of one refers to the situation at the start of our sample period and zero
        means that all NPIs included in the index are turned on. The other lines in
        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect of
        contact reductions, increased hygiene regulations, and seasonality. See
        Appendix~\ref{subsec:policies} for separate plots of the three factors by contact
        type.}
\end{figure}

By March 2021, the set of policy instruments had become much more diverse. Around the
turn of the year, the first people were vaccinated with a focus on older age groups and
medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). Until the end of May,
43\% had received at least one dose of a vaccine. In late 2020, rapid tests started to
replace regular PCR tests for staff in many medical and nursing facilities. These had to
be administered by medical doctors or in pharmacies. At-home tests approved by
authorities became available in mid-March, rapid test centers were opened, and one test
per person and week was made available free of charge. In several states, customers were
only allowed to enter certain stores with a recent negative rapid test result. These
developments are characteristic of many countries: The initial focus on NPIs to slow the
spread of the disease has been accompanied by vaccines and a growing acceptance and use
of rapid tests. At broadly similar points in time, novel strains of the virus have
started to pose additional challenges.

We draw simulated samples of agents from the population structure in September 2020 and
use the model to predict recorded infection rates until the end of May 2021. See
Supplementary Material sections~\ref{subsec:synthetic_population} and
\ref{sub:initial_conditions} for details. The blue line in
Figure~\ref{fig:aggregated_fit} shows that our model's predictions are very close to
officially recorded cases in the aggregate. This is also true for infections by age and
geographical region, which are shown in the supplementary materials
(Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).

The effects of various mechanisms can be disentangled due to the distinct temporal
variation in the drivers of the pandemic. Next to the stringency index, the three lines
in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions,
increased hygiene regulations, and seasonality evolved since early September for each of
the three broad contact networks. For example, a value of 0.75 for the work multiplier
means that if the environment was the same as in September (levels of infection rates,
no rapid tests or vaccinations, only the wildtype virus present), infections at the
workplace would be reduced by 25\%. The lines show the product of the effect of contact
reductions, increased hygiene regulations, and seasonality. Two aspects are particularly
interesting. First, all lines broadly follow the stringency index and they would do so
even more if we left out seasonality and school vacations (roughly the last two weeks of
October, two weeks each around Christmas and Easter, and some days in late May). Second,
the most stringent regulations are associated with the period of strong decreases in new
infections between late December 2020 and mid-February 2021. The reversal of the trend
is associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
during May 2021 is associated with at least weekly rapid tests for 42\% of the
population, a vaccination rate that rose from 28\% to 43\%, and seasonality lowering the
relative infectiousness of contacts, most strongly for other contacts.

In order to better understand the contributions of rapid tests, vaccinations, and
seasonality on the evolution of infections in 2021, Figure~\ref{fig:2021_scenarios_broad}
considers various scenarios. NPIs are always the same as in the baseline scenario.
Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
and three scenarios turning each of these factors on individually.
Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
\citep{Shapley2016} to decompose the difference in total infections between the scenario
without any of the three factors and our main specification.

\begin{figure}[!tp]
    \centering

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_new_known_case}
        \caption{{Recorded cases: 2021 scenarios}}
        \label{fig:2021_scenarios_recorded}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/effect_of_channels_on_pessimistic_scenario/full_newly_infected}
        \caption{{Total cases: 2021 scenarios}}
        \label{fig:2021_scenarios_newly_infected}
    \end{subfigure}

    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_channels_area}
        \caption{Decomposition of the difference between the scenario without any of the
            three factors and the main scenario in
            Figure~\ref{fig:2021_scenarios_newly_infected}.}
        \label{fig:2021_scenarios_decomposition}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/full_decomposition_rapid_tests_area}
        \caption{Decomposition of the difference between the scenario without rapid
            tests and the main scenario in
            Figure~\ref{fig:2021_scenarios_newly_infected}.}
        \label{fig:2021_scenarios_decomposition_tests}
    \end{subfigure}

    \caption{The effect of different interventions on recorded and actual infections}
    \label{fig:2021_scenarios_broad}

    \floatfoot{\noindent \textit{Note:} The blue line in
        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
        variant is introduced in the same way; vaccinations, rapid tests, and seasonality
        remain at their January levels. The other scenarios turn each of these three
        factors on individually. The decompositions in
        Figures~\ref{fig:2021_scenarios_decomposition} and
        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
        are explained more thoroughly in Appendix~\ref{sub:shapley_value}}
\end{figure}

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active, the
final level is always the highest in case of the vaccination campaign (orange lines). The
Shapley value decomposition shows that vaccinations contribute 16\% to the cumulative
difference between scenarios. Reasons for this are the slow start---it took until
March~24th until 10\% of the population had received their first vaccination, the 20\%
mark was reached on April 19th---and the focus on older individuals. These groups
contribute less to the spread of the disease than others due to a lower number of
contacts (see Supplementary
Material~\ref{subsec:assortativity}).\comment[id=K]{Assortativity is the wrong section
here. They have few contacts in our model because they tend to live in small households
and don't have work or educ contacts. They are also more likely to meet other old people
but that's a different channel from what is written in the text.} By the end of our study
period, when first-dose vaccination rates reached around 40\% of the population, the
numbers of new cases would have started to decline. It is important to note that the
initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
fatality was rate considerably lower during the third wave when compared to the second
(4.4\% between October and February and 1.4\% between March and the end of
May).\comment[id=K]{This needs a source.}

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and recorded cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace \citep[this is in line with results in][, which our seasonality measure
is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
quantitatively important factor determining the evolution of the pandemic, explaining
most of the early changes and 43\% of the cumulative difference by the end of May.

A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
is crucial to differentiate between recorded cases and actual cases. Additional testing
means that additional infections will be recorded which would otherwise remain
undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large.
Until late April, recorded cases are higher in the scenario with rapid testing alone when
compared to the setting where none of the three mechanisms are turned on. The effect on
total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
population performed weekly rapid tests in March on average, new infections on April 1
would have been reduced by 53\% relative to the scenario without vaccinations, rapid
tests, or seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests only and the main specification into the three channels for
rapid tests. Tests at schools have the smallest effect, which is largely explained by
schools not operating at full capacity during our period of study and the relatively
small number of students.\footnote{18\% of our population are in the education sector
(pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
come from tests at the workplace. Despite the fact that rapid tests for private reasons
are phased in quite late, they make up for more than half of the total effect. The reason
lies in the fact that a substantial share of these tests is driven by an elevated
probability to carry the virus, i.e., showing symptoms of CoViD-19 or following up on a
positive test of a household member. The latter is essentially a form of contact tracing,
which has been shown to be very effective \citep{Contreras2021,
Kretzschmar2020,Fetzer2020}.

Two of the most contentious NPIs concern schools and mandates to work from
home.\comment[id=K]{We should decide if we want to keep this section or prefer to show
some rapid test statistics.} In many countries, schools switched to remote instruction
during the first wave, so did Germany. After the summer break, they were operating at
full capacity with increased hygiene measures, before being closed again from
mid-December onward. Some states started opening them gradually in late February, but
operation at normal capacity did not resume until the beginning of June.
Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
schools starting at Easter, at which point rapid tests had become widely available. We
estimate the realized scenario to have essentially the same effect as a situation with
closed schools. Under fully opened schools with mandatory tests, total infections would
have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}. \comment[id=HM]{Need
a back-of the envelope calculation} To use another metric, the effective weekly
reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021}---and in particular on those with low socio-economic status---these results
in conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when rapid test
are not available at scale, trade-offs may well be different.

\begin{figure}[!tp]
    \centering

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
        \caption{{Effects of different schooling scenarios}}
        \label{fig:school_scenarios}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
        \caption{{Effects of different work scenarios}}
        \label{fig:workplace_scenarios}
    \end{subfigure}
    \vskip3ex

    \caption{Effects of different scenarios for policies regarding schools and workplaces.}
    \label{fig:school_workplace_scenarios}

    \floatfoot{\noindent \textit{Note:} Blue lines in both figures refer to our baseline
        scenario; they are the same as in
        Figure~\ref{fig:2021_scenarios_newly_infected}. Interventions start at Easter
        because there were no capacity constraints for rapid tests afterwards.}

\end{figure}

Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
receiving tests, testing at the workplace has larger effects than mandating employees to
work from home. Whether the share of workers working at the usual workplace is reduced
or increased by ten percent changes infection rates by 2.5\% or less in either
direction. Making testing mandatory twice a week---assuming independent compliance by
employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
rapid tests offers by employers to the level of March would have increased infections by
13\%.

Our analysis has shown that during the transition to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
estimated to have cost around 20~Euros per capita \citep{Wollmershauser2021}; retail
prices for rapid tests were below one Euro in early June 2021. Despite these large
effects, the results on testing likely understate the benefits. Disadvantaged groups are
less likely to be reached by testing campaigns relying on voluntary participation
\citep[e.g.][]{StillmanTonin2021}; at the same time, these groups have a higher risk to
contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the
workplace will extend more into these groups. The same goes for individuals who exhibit
a low level of compliance with CoViD-19-related regulations. Compared to vaccinations,
rapid testing programmes allow a much quicker roll-out, making it arguably the most
effective tool to contain the pandemic in the short run.