### 2021-07-12 src/paper/appendix_results.tex
Use greek letters in medical parameters section.

\section{Additional Results}
\label{sec:additional_results}

\input{paper/results/fit}

\input{paper/results/validation}

\input{paper/results/share_known_cases}

\input{paper/results/rapid_tests_statistics}

\input{paper/results/scenarios}

\input{paper/results/comparison_with_other_studies}