### 2021-07-12 src/paper/results/validation.tex

\subsection{Model Validation}
\label{subsec:model_validation}

Achieving a good in-sample fit does not necessarily guarantee that our model will also be
able to make out of sample predictions. For example, it could be that the results are
very sensitive to the exact number of vaccinations, the work mobility multiplier
($\rho_{w,\:attend,\:t}$) or the number of rapid tests (governed by the $\pi$ parameters)
that are performed -- all of which are things that cannot be exactly known ex-ante.

In this section we compare simulated infections that use all available data with
out of sample predictions that only use data that was available at March 1 2021.

For the out of sample predictions we predict the number of vaccinations between March and
June with a simple linear regression model that was fitted on vaccine data from February.
This prediction model is pessimistic compared to the actual number of vaccinations. The
work mobility multiplier ($\rho_{w,\:attend,\:t}$) is predicted to be constant at a value
of 0.75, which is an approximate average of the second half of February. This turned out
to be optimistic.

The area that is fraught with the most uncertainty is the introduction of rapid tests,
because it comprises both supply and demand factors. Moreover, accurately predicting the
number of rapid tests is expected to be important because rapid tests play a large role
for the transmission dynamic.

We therefore make a scenario analysis with different assumptions on the availability of
rapid tests. The number of rapid tests performed in each scenario can be seen in
Figure~\ref{fig:robustness_check_rapid_test_params}. All scenarios are the same until
March 1 and have the same level of rapid tests when all supply constraints are resolved.
They differ in the date at which the full number of tests is reached. For students
($\pi_{students,\:t}$) and teachers ($\pi_{teacher,\:t}$) the full number of rapid tests
is reached after the Easter holidays in all scenarios. For rapid tests in the workplace
($\pi_{w,\:s,\:t}$) and private rapid tests ($\pi_{private,\:t}$) it is reached between
May 1 and June 10, depending on the scenario.

\begin{figure}[ht] % Robustness Check Rapid Test Fade In Params
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand/robustness_check_params_early_shares}
    \caption{Rapid Test Parameters: Early Scenario}
    \label{fig:robustness_early_params}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand/robustness_check_params_medium_shares}
    \caption{Rapid Test Parameters: Medium Scenario}
    \label{fig:robustness_medium_params}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{figures/results/figures/data/testing/rapid_test_demand/robustness_check_params_late_shares}
    \caption{Rapid Test Parameters: Late Scenario}
    \label{fig:robustness_late_params}
  \end{subfigure}

  \caption{Rapid Test Introduction in the Three Scenarios}
  \label{fig:robustness_check_rapid_test_params}

  \floatfoot{\noindent \textit{Note:} Number of rapid tests performed in the different
  prediction scenarios. All scenarios are the same until March 1 and have the same level
  of rapid tests when all supply constraints are resolved. They differ in the date at
  which the full number of tests is reached. For students ($\pi_{students,\:t}$) and
  teachers ($\pi_{teacher,\:t}$) the full number of rapid tests is reached after the
  Easter holidays in all scenarios. For rapid tests in the workplace ($\pi_{w,\:s,\:t}$)
  and private rapid tests ($\pi_{private,\:t}$) it is reached between May 1 and June 10,
  depending on the scenario. }
\end{figure}

Moreover, the out of sample predictions assume that the share of detected cases
($\psi_t$) that would have been obtained without rapid tests is not affected by the
Easter holidays because the extent to which this was the case was estimated from case
numbers in April.

The results of the out of sample prediction are displayed in
Figure~\ref{fig:robustness_check_detailed}. While all scenarios considerably deviate from
the ex-post scenario, they all reproduce the steep increase of cases until the end of
April, followed by a decline until June. We can therefore conclude that our main results
are not sensitive to measurement errors in the number of rapid tests, vaccinations or
mobility data.

\begin{figure}[ht] % Robustness Check
  \centering
  \begin{subfigure}[b]{.425\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_new_known_case}
    \caption{Reported Cases}
    \label{fig:robustness_check_new_known_case}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[b]{.425\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/results/figures/scenario_comparisons/robustness_check/full_newly_infected}
    \caption{Total Cases}
    \label{fig:robustness_check_newly_infected}
  \end{subfigure}
  \caption{Out of Sample Prediction for Reported and Total Cases from March to June
  2021.}
  \label{fig:robustness_check_detailed}
  \floatfoot{\noindent \textit{Note:} The ex-post scenario is an in-sample prediction
  that uses all available information and is very close to actual case numbers. For the
  other scenarios data on vaccinations, work mobility and rapid tests that became
  available after March 1 have been replaced by prediction models that are calibrated
  with data from February. Moreover, they do not model a lower number of detected cases
  over the Easter holidays. The different scenarios make different assumptions on the
  date at which full availability of rapid tests is reached. While the out of sample
  predictions differ substantially for the exact case numbers at the beginning of June
  (between 20 and 70 cases per million), they can all reproduce the decline in case
  numbers that is jointly driven by seasonality, large scale rapid tests and
  vaccinations.}
\end{figure}

\FloatBarrier