### 2021-07-12 src/paper.tex
Add science style.

\pdfoutput=1

% To ``accept'' all changes in the document:
% \PassOptionsToPackage{final}{changes}

\documentclass[a4paper, 10pt]{article}
% fleqn: flush-left aligned equations

\newcommand{\papertitle}{
The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs
}

\input{templates/Preamble_Fundamental_packages}
\hypersetup{
    pdftitle={\papertitle},
    pdfauthor={Janoś Gabler, Tobias Raabe, Klara Röhrl, Hans-Martin von Gaudecker}
}

% packages ======================================================================

\input{templates/Preamble_Page_layout_general}
\input{templates/Preamble_Page_layout_paper}
\subimport{templates/}{Preamble_Fonts_Charter_FiraSans}
\input{templates/Preamble_Sansserif_math}
% \input{templates/Preamble_BibLaTeX_AER}
% \input{templates/Preamble_BibLaTeX_APA}
\input{templates/Preamble_BibLaTeX_Science}
\input{templates/Preamble_Commenting_and_highlighting}
\input{templates/Preamble_Miscellaneous}

% Add global hyphenation rules to avoid unwelcome linebreaks:
% \hyphenation{Bu-shong Della-Vigna}

\bibliography{references.bib}

% \usepackage[newfloat, outputdir=../bld, cache=false]{minted}

% Define environments to comment on the sidebar of the PDF.
\definechangesauthor[name={Tobias}, color=alizarin]{T}
\definechangesauthor[name={Janos}, color=internationalorange]{J}
\definechangesauthor[name={Klara}, color=dartmouthgreen]{K}
\definechangesauthor[name={HM}, color=blue]{HM}
% Use \replaced[id=T]{<to-be-replaced>}{<new>} or the same with \deleted to propose
% changes.

\begin{document}

\selectlanguage{USenglish}
\frenchspacing

\input{paper/titlepage}

\maketitle
\pagenumbering{arabic}
\setcounter{page}{1}

\newpage

\input{paper/report}

\phantomsection % Needed for the bibliography bookmark to point to the correct page

\begin{refcontext}[sorting=nyt]  % Sort BIBLIOGRAPHY by alphabet (while CITATIONS are sorted by year)
\printbibliography[heading=bibintoc]
\end{refcontext}

\clearpage

\begin{center}
    {
        \Large 
        Supplementary Material for\\[2ex]

        \textbf{\papertitle}
    }
    \vskip4ex
    {
        \large
        Janoś Gabler,

        Tobias Raabe,

        Klara Röhrl,

        Hans-Martin von Gaudecker

    }
\end{center}

\clearpage

\begin{appendices}
    \begin{refsection}

    This paper aims to be completely reproducible. The code for the general model can be
    found at
    \href{https://github.com/covid-19-impact-lab/sid/}{https://github.com/covid-19-impact-lab/sid/}
    and its documentation at
    \href{https://sid-dev.readthedocs.io/en/latest/}{https://sid-dev.readthedocs.io/en/latest/}.
    The code for the application to Germany can be found at
    \href{https://github.com/covid-19-impact-lab/sid-germany/}{https://github.com/covid-19-impact-lab/sid-germany/}
    with a very rudimentary documentation at
    \href{https://sid-germany.readthedocs.io/en/latest/}{https://sid-germany.readthedocs.io/en/latest/}.

    \input{paper/appendix_data}

    \FloatBarrier

    \input{paper/appendix_model}

    \input{paper/appendix_results}

    \printbibliography[heading=bibintoc]

    \end{refsection}
\end{appendices}

\end{document}