### 2021-07-13 src/paper/model/code.tex

The code used for this paper is open source and available under the MIT License. It is
split into two parts
\begin{itemize}
    \item The code for the model can be found at
   \href{https://github.com/covid-19-impact-lab/sid/}{https://github.com/covid-19-impact-lab/sid/}
   and its documentation at
   \href{https://sid-dev.readthedocs.io}{https://sid-dev.readthedocs.io}.
   \item The code for the application to Germany can be found at
   \href{https://github.com/covid-19-impact-lab/sid-germany/}{https://github.com/covid-19-impact-lab/sid-germany/}
   with a shorter documentation at
   \href{https://sid-germany.readthedocs.io}{https://sid-germany.readthedocs.io}.
\end{itemize}