### 2021-07-13 src/paper/model/param_tables.tex
Shortened some parts of main text, added discussion of rapid tsts.

\clearpage

\begin{landscape}

\subsection{Overview Model Parameters} % (fold)
\label{sub:param_tables}

\comment[id=HM]{Copied from main text. Do we describe the parameter for fraction of
vaccination-induced immunity/sterility  somewhere? I could not find it. Nothing special
about the 75\%...

Also, start out more positively. Something like ``We do not distinguish between first
and second doses for computational reasons. We thus use a weighted average ...''

75\% is lower than what is usually reported for after the second dose
of the Biontech/Pfizer vaccine, which is most commonly used in Germany. We choose it
because our model neither includes booster shots, nor does it allow vaccinated
individuals who became immune to transmit the disease\citep{Petter2021,
LevineTiefenbrun2021, Pritchard2021}. If anything, these assumptions would overstate the
effect of vaccines for our study period. This would be different if a large fraction of
vaccinated individuals had received a second dose already.

}

\FloatBarrier

\begin{table}[htb]
\centering
\caption{Contacts, Matching and Policies}
\label{tab:contacts_and_matching}
\input{../bld/param_tables/contacts_matching_policies}
\end{table}

\begin{table}[htb]
\centering
\caption{Infection Probabilities and Virus Variants}
\label{tab:infection_probs}
\input{../bld/param_tables/infection_probs}
\end{table}

\begin{table}[htb]
\centering
\caption{Disease and Vaccination Model}
\label{tab:disease_and_vacc_model}
\input{../bld/param_tables/disease_model}
\end{table}

\begin{table}[htb]
\centering
\caption{Rapid Testing}
\label{tab:rapid_testing}
\input{../bld/param_tables/rapid_testing}
\end{table}

\begin{table}[htb]
\centering
\caption{PCR Testing, Case Detection and Behavioral Response}
\label{tab:pcr_testing_and_case_detection}
\input{../bld/param_tables/pcr_testing_and_case_detection}
\end{table}

\FloatBarrier

\end{landscape}

\FloatBarrier