### 2021-07-14 src/paper/model/policies.tex

\subsection{Reducing Numbers of Contacts via NPIs}
\label{sec:policies}

 \comment[id=K]{I think we need to make sure we distinguish very clearly what is possible
 and what's implemented.}

Our model makes it very easy to model a wide range of NPIs, either in isolation or
simultaneously. This is important for two reasons: Firstly, it allows to predict and
quantify the effect of novel NPIs. Secondly, it allows to model the actually implemented
policy environment in great detail, which is necessary to use the full time series
of infections and fatality rates to estimate the model parameters.\footnote{
See \citet{Avery2020} for an explanation why it can be harmful to use too long time
series to estimate simple SEIR type models.}

Instead of thinking of policies as completely replacing how many contacts people have,
it is often more helpful to think of them as adjusting the pre-pandemic number of
contacts. Therefore, we implement policies as a step that happens after the number of contacts is
calculated but before individuals are matched.

On an abstract level, a policy is a functions that modifies the number of contacts of
one contact type. This function can be random or deterministic. For example, school
closures simply set all school contacts to zero. A work from home mandate leads to a
share of workers staying home every day whereas those who cannot work from home are
unaffected. Hygiene measures at work randomly reduce the number of infectious contacts
for all workers who still go to work.

Policies can also interact. For example, school vacations are temporally reducing school
contacts to zero while at the same time increasing other contacts to account for
increased leisure activities and family visits during this time. This is important to
reproduce the finding that school vacations do not reduce infection numbers even though
schools lead to infections when open \citep{Isphording2021}.

The most complex policies are typically found in the education sector. Since the
beginning of 2021 schools have switched back and fourth between full closures, split
class approaches with alternating schedules for some or all age groups and reopening
while maintaining hygiene measures. On top of that there are different policies for
allowing young students whose parents work full time to attend school even on days where
they normally would not. For details on how we calibrate these policies see
Section~\ref{subsec:policies}.

Importantly, policies can depend on the health states of participating individuals. For
example children rarely go to school when they have symptoms. It would even be possible
to quarantine entire school classes if one student tested positive and many other forms
of contact tracing. For an application of our model showcasing private contact tracing in
the context of the Christmas holidays see \cite{Gabler2020}.

Not all things that reduce contacts compared to the pre-pandemic level are driven by
NPIs. Therefore, we also model endogenous contact reductions that depend on the health
state of individuals. Other possible factors could include things such as the local
incidence. The extent to which contacts are reduced can be calibrated from surveys.