### 2021-07-15 src/paper/data/course_of_disease.tex
Add values and refer to sections in tables B.2 until B.4. Add paragraph on immunity and infection probability in the course of disease section.

\subsection{Course of Disease}
\label{sub:data_course_of_disease}

This section discusses the parameters governing the course of disease, their sources and
how we arrived at the distributions used in the paper.
See Figure~\ref{fig:model_disease_progression} for a summary of our
disease progression model.

% Susceptibility and Infectiousness
 \comment[id=K]{@Janos: read this.}
The first stage of any disease is the infection. As detailed in
Equation~\ref{eq:infection_probability_formula} the infection
probability depends on the contact type, the calendar date to determine the seasonality,
the age group of the susceptible person and the variant the infectious person is
carrying.
%
The base infection probability of each contact type ($\beta_c$) is estimated inside our
model (Section~\ref{subsec:estimated_params}). How we model and set the seasonality is
detailed in Section~\ref{subsec:seasonality}. For the susceptibility of each age group
($\zeta_a$) we take the estimates of \citet{Davies2020} (Extended Data Fig. 4). Lastly,
we calibrate the infectiousness of the B.1.1.7 variant ($\sigma_{B.1.1.7}$) from
\citet{Davies2021} to 1.67.

% Duration until Infectiousness and Symptoms ----------------------------------

% Latency Period
We denote the latent period---i.e., the time span between infection and the start of
infectiousness---by $\gamma_{infectious}$. \cite{Zhao2021} estimate the latent period to
last 3.3 days (95\% CI: 0.2, 7.9) on average. In line with this estimate our latent
period lasts one to five days.

% Asymptomatic Share
Once individuals become infectious, a share of them goes on to develop symptoms while
others remain asymptomatic. We rely on data by \cite{Davies2020} for the age-dependent
probability to develop symptoms. It varies from 25\% for children and young adults to
nearly 70\% for the elderly.
% Length of the presymptomatic stage
Similar to \citet{Peak2020} and in line with \citet{He2020} we set the length of the
presymptomatic stage of age group $a$, $\gamma_{symptoms,\:a}$ to be one or two days. The
probability to become symptomatic for age group $a$ is split equally between one and two
days. This combined with our latency period leads to an incubation period that is in line
with the meta analysis by \citet{McAloon2020}.

% Duration of Infectiousness ----------------------------------------------------------

We assume that the duration of infectiousness ($\gamma_{stop\:infectious}$) is the same
for both symptomatic and asymptomatic individuals as evidence suggests little differences
in the transmission rates between symptomatic and asymptomatic patients (\citet{Yin2020})
and that the viral load between symptomatic and asymptomatic individuals are similar
(\citet{Zou2020}, \citet{Byrne2020}, \citet{Singanayagam2020}). Our distribution of the
duration of infectiousness is based on \citet{Byrne2020}. For symptomatic cases they
arrive at zero to five days before symptom onset (see their figure 2) and three to eight
days of infectiousness afterwards.\footnote{Viral loads may be detected much later but
eight days seems to be the time after which most people are culture negative, as also
reported by \citet{Singanayagam2020}.} Excluding the most extreme combinations, we arrive
at 3 to 11 days as the duration of infectiousness.

% Duration of Symptoms ----------------------------------------------------------

We use the duration to recovery of mild and moderate cases reported by \cite[Figure~S3,
Panel~2]{Bi2020} for the duration of symptoms for non-ICU requiring symptomatic cases
($\gamma_{stop\:symptoms}$). We only disaggregate by age how likely individuals
are to require intensive care. %

For the time from symptom onset until need for intensive care we rely on data by
\cite{Stokes2020}) and \cite{Hinch2021} ($\gamma_{icu,\:a}$). For those who will require
intensive care we follow \citet{Chen2020} who estimate the time from symptom onset to ICU
admission as $8.5 \pm 4$ days. This aligns well with numbers reported for the time from
first symptoms to hospitalization: \citet{Gaythorpe2020} report a mean of 5.76 with a
standard deviation of four. We assume that the time between symptom onset and ICU takes
four, six, eight or ten days with equal probabilities.

% Death and Recovery from ICU ----------------------------------------------------------
We take the survival probabilities and time to death and time until recovery
($\gamma_{stop\:icu\:a}$ and $\gamma_{dead,\:a}$) from intensive care from
\citet{Hinch2021}. They report time until death to have a mean of 11.74 days and a
standard deviation of 8.79 days. To match this we discretize that 41\% of individuals who
will die from Covid-19 do so after one day in intensive care, 22\% die after twelve days,
29\% after 20 days and 7\% after 32 days. Again, we rescale this for every age group
among those that will not survive. For survivors \cite{Hinch2020} reports a mean duration
of 18.8 days until recovery and a standard deviation of 12.21 days. We discretize this
such that of those who recover in intensive care, 22\% do so after one day, 30\% after 15
days, 28\% after 25 days and 18\% after 45 days.

% Immunity and Vaccination -------------------------------------------------------------

 \comment[id=K]{@Janos: Read this!}

Individuals can become immune either through infection ($\gamma_{immune}$) or vaccination
($\gamma_{vacc,\:d}$). As reinfections are estimated to be extremely rare
\citep{AbuRaddad2020}, we simply set the immunity period to one year with probability
one, i.e. everyone that has been infected enjoys immunity for the rest of the simulation
period.

% vaccinations
The second route to immunity is through vaccination. Germany has mostly relied on the
Pfizer-BioNTech BNT162b2 and Oxford-AstraZeneca ChAdOx1-S vaccines with smaller shares of
the Moderna and Johnson\&Johnson vaccines \citep{Impfdashboard}. As \citet{Pritchard2021}
and \citet{Harris2021} find no difference in the effectiveness between the two most
common vaccines, we do not distinguish between vaccine types.

Immunity is binary in our model, i.e. individuals become either sterile immune or remain
susceptible. Thus, we cannot simply use the reported effectiveness but must also include
the risk of asymptomatic and sub-clinical reinfection among the vaccinated in our
probability to become immune upon vaccination. This is important as there is by now ample
evidence that vaccinated individuals can still get infected with SARS-CoV-2 and transmit
the disease \citep{Petter2021, LevineTiefenbrun2021, Harris2021}.

The reported effectiveness for BNT162b2 is estimated to be 90\% 21 days after the first
shot \citep{Hunter2021}. The effectiveness does not increase much through the booster
shot as \citet{Thompson2021} report 90\% (95\% CI = 68\%–97\%) effectiveness against
PCR-confirmed infections after two doses for mRNA vaccines in general. We therefore do
not distinguish between the first and the booster shot.

On the other hand, \citet{Lipsitch2021} report a lower bound on transmission for
the very similar Moderna vaccine of 61\%.

Taking this evidence together we simplify to 35\% of individuals developing immunity 14
days after their vaccination and 40\% after 21 days. For the remaining 25\% we assume
that the vaccine does not protect them. %
This may seem slightly optimistic for how early vaccinations may already have an effect
when looking at infection rates like \citet{Hunter2021}. However, it can be expected that
individuals underestimate the time it takes for the vaccine to work and behave less
carefully after receiving a vaccination. We err on the side of being too optimistic for
the effectiveness of vaccine and assume a significant protection after 14 days as
\citet{LevineTiefenbrun2021} report strongly reduced viral loads already 12 to 28 days
after the first vaccine.

\FloatBarrier