### 2021-07-15 src/paper/appendix_data.tex
Add section for pcr testing, case detection and behavioral response. Finishes adding a paragraph or sentence for every parameter in the tables of B.12.

\section{Data and Parameters}
\label{sec:data_and_parameters}

The model is described by a large number of parameters that govern the number of contacts
a person has, the reduction in contacts due to NPIs, the demand for rapid tests and PCR
tests, the likelihood of becoming infected on each contact, the likelihood of developing
light or strong symptoms or even dying from the disease as well as the duration each
stage of the disease takes.

\input{paper/data/course_of_disease}

\input{paper/data/synthetic_population}

\input{paper/data/number_of_contacts}

\input{paper/data/assortativity}

\input{paper/data/policies}

\input{paper/data/rapid_test_demand}

\input{paper/data/share_known_cases}

\input{paper/data/testing_and_behavioral_response}