### 2021-07-15 src/paper/model/code.tex

\subsection{Reproducibility}
\label{subsec:code}

The source code used for this paper is open source and available under the MIT License.
It is split into two parts
\begin{itemize}
    \item The source code for the model can be found at
   \href{https://github.com/covid-19-impact-lab/sid/}{https://github.com/covid-19-impact-lab/sid/}
   and its documentation at
   \href{https://sid-dev.readthedocs.io}{https://sid-dev.readthedocs.io}.
   \item The source code for the application to Germany can be found at
   \href{https://github.com/covid-19-impact-lab/sid-germany/}{https://github.com/covid-19-impact-lab/sid-germany/}
   with a shorter documentation at
   \href{https://sid-germany.readthedocs.io}{https://sid-germany.readthedocs.io}.
\end{itemize}

We are grateful to the authors and contributors of the following software packages upon
which our software is built: conda \cite{Anaconda2016}, conda-forge
\cite{CondaForge2015} dask \cite{Rocklin2015}, estimagic \cite{Gabler2020a}, holoviews
\cite{Stevens2015}, matplotlib \cite{Hunter2007}, numba \cite{Lam2015}, numpy
\cite{Harris2020}, pandas \cite{Reback2020,McKinney2010}, pytask \cite{Raabe2020},
Python \cite{VanRossum1995}, scipy \cite{Virtanen2020}.