### 2021-07-15 src/figures/results/tables/student_infections_table.tex

\begin{tabular}{lr}
\toprule
{} &  predicted total infections among 5-14 year olds from Easter until 2021-05-31 \\
scenario                               &                                                                               \\
\midrule
 educ open after easter  without tests &                                             951185 \\
 educ open after easter  with tests    &                                             722502 \\
 close educ after easter               &                                             559526 \\
 baseline                              &                                             582466 \\
\bottomrule
\end{tabular}