### 2021-07-15 src/paper/model/initial_conditions.tex

\subsection{Initial Conditions} % (fold)
\label{sub:initial_conditions}

Consider a situation where you want to start a simulation with the beginning set amidst
the pandemic. It means that several thousands of individuals should already have
recovered from the disease, be infectious, symptomatic or in intensive care at the start
of your simulation. Additionally, the sample of infectious people who will determine the
course of the pandemic in the following periods is likely not representative of the whole
population because of differences in behavior (number of contacts, assortativity), past
policies (school closures), etc.. The distribution of health states in the population at
the beginning of the simulation is called initial conditions.

To come up with realistic initial conditions, we match reported infections from official
data to simulated individuals by age group and county. We use one month of data to
generate initial conditions with in all possible health states. Meanwhile health states
evolve until the beginning of the simulation period without simulating infections by
contacts. We also correct reported infections for a reporting lag and scale them up by
the share of detected cases to arrive at the true number of infections.

% subsection initial_conditions (end)