### 2021-07-16 src/science/cover_letter.tex
Tiny updates to cover letter.

\documentclass[a4paper, 10pt]{article}

\title{The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs}

\usepackage{url}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{Janoś Gabler,$^{1, 2}$\\Tobias Raabe,$^{3}$\\Klara Röhrl,$^{1}$\\Hans-Martin von Gaudecker,$^{2,4\ast}$\\
\\
\normalsize{$^{1}$Bonn Graduate School of Economics}\\
\normalsize{$^{2}$IZA Institute of Labor Economics}\\
\normalsize{$^{3}$Private sector}\\
\normalsize{$^{4}$Rheinische Friedrich-Wilhelms-Universität Bonn}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  hmgaudecker@uni-bonn.de.}
}

\begin{document}

\maketitle

\noindent
\textbf{Main point}
\vskip1ex

\noindent
Our main methodological contribution is to calibrate a very detailed and flexible agent
based model to a long time series of data during the CoVid-19 pandemic. Our core
substantive result is the quantitative importance of rapid testing as a supplement to
NPIs. Due to the slow vaccine roll-out in most countries, this will remain topical for
some time.

\vskip2ex

\noindent
There were no submissions to other journals, nor discussions with editors prior to
submitting the report.

\vskip2ex

\noindent
A draft of this report was reviewed by Prof. Jörg Stoye and Prof. Philipp Eisenhauer.

\vskip2ex

\noindent
A draft of this report has been published as a pre-print on arXiv
(\url{https://arxiv.org/abs/2106.11129}). It is not under consideration for publication
elsewhere.

\vskip2ex

\noindent
All data is freely available in linked to in the supplementary material. Only one data
set, the microcensus of 2010, provided by the FDZ (Forschungsdatenzentren) requires
registration prior to access
(\url{http://www.forschungsdatenzentrum.de/de/campus-files})

\vskip2ex

\noindent
The submitted materials include sections on materials and methods and supplementary
text.

\end{document}