### 2021-09-29 src/paper/data/rapid_test_demand.tex
Add additional references by HiWis to the rapid test demand section.

\subsection{Rapid Test Demand}
\label{subsec:rapid_test_demand}

In our model, there are five reasons why rapid tests are done:

\begin{enumerate}
    \item someone plans to have work contacts
    \item someone is an employee or student of an educational facility
    \item a household member has tested positive or developed symptoms
    \item someone has developed symptoms but has not received a PCR test
    \item someone plans to participate in a weekly non-work meeting
\end{enumerate}

%%% \subsubsection{work rapid tests}

For work contacts, we know from the COSMO study (\cite{Betsch2021}, 20th/21st of April)
that 60\% of workers who receive a test offer by their employer regularly use it
($\pi_{w,\:d}$). We assume this share to be time constant.

In addition, there are some surveys that allow us to trace the expansion of employers
who offer tests to their employees ($\pi_{w,\:s,\:t}$). Mid march, 20\% of employers
offered tests to their employees \citep{DIHK2021}. In the second half of March, 23\% of
employees reported being offered weekly rapid tests by their employer
\citep{Ahlers2021}\footnote{\cite{IZA2021b} reports a somewhat higher share of 35\%
around March 20}. This share increased to 61\% until the first days of April
\citep{IZA2021c, IZA2021}. Until mid April 72\% of workers were expected to receive a
weekly test offer \citep{IZA2021c, IZA2021}. However, according to surveys conducted in
mid April \citep{Betsch2021}, less than two thirds of individuals with work contacts
receive a test offer.\footnote{\cite{IZA2021d} report 80\% in mid April which is about
12\% points less than what we arrive at.} Starting on April 19th employers were required
by law to provide two weekly tests to their employees \citep{Bundesanzeiger2021}.
However, compliance was very incomplete and only reached 80\% \citep{IAB2021b,
Ahlers2021b}. We interpolate between these points linearly, arriving at
Figure~\ref{fig:work_rapid_test_offers}. In addition, we increase the frequency of
testing ($\theta_{t,\:work}$) from weekly to twice weekly during April.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/share_of_workers_with_rapid_test_offer_at_work}
    \caption{Share of Individuals Being Offered a Rapid Test at Work.}
    \floatfoot{\noindent \textit{Note:} Employees were never subject to mandatory tests.
    According to surveys only 60\% of workers regularly used rapid test offers by their
    employers (\cite{Betsch2021}, 20th/21st of April).}
\end{figure}
\FloatBarrier

%%\subsubsection{educ rapid tests}

We assume that employees in educational facilities start getting tested in 2021 and that
by March 1st 30\% of them ($\pi_{teacher,\:t}$) are tested weekly
($\theta_{before\:Easter,\:educ} = 7$). The share increases to 90\% for the week before
Easter. At that time both Bavaria \citep{STMGP2021} and Baden-Württemberg
\citep{MinisteriumKultus2021} were offering tests to teachers and North-Rhine Westphalia
\citep{SchulministeriumNRW2021} and Lower Saxony \citep{NSMK2021} were already testing
students and tests for students and teachers were already mandatory in Saxony
\citep{SMK2021}. After Easter we assume that 95\% of teachers get tested twice per week
($\theta_{after\:Easter,\:educ} = 3$).

Tests for students started later \citep{MinisteriumKultus2021, SchulministeriumNRW2021}
so we assume that they only start in February and only 10\% of students get tested by
March 1st ($\pi_{students,\:t}$). Relying on the same sources as above we approximate
that by the week before Easter this share had increased to 40\%
\citep{SchulministeriumNRW2021}. After Easter the share of students receiving twice
weekly tests is set to 75\%. This is based on tests becoming mandatory in Bavaria
\citep{BayerischeStaatskanzlei2021} and North Rhine-Westphalia
\citep{SchulministeriumNRW2021b} after their Easter breaks and on the 19th in
Baden-Württemberg \citep{KMBaWue2021}, after which we assume twice weekly rapid tests to
be mandatory for all students in Germany. Again, we interpolate linearly between these
points and arrive at the purple line for teachers and the red line for school students in
Figure~\ref{fig:rapid_test_demand}.

%%\subsubsection{private rapid tests}

To limit our degrees of freedom, we only have one parameter that governs how many
individuals do a rapid test because of any of the private demand reasons
($\pi_{private,\:t}$).\footnote{The reasons that can lead to an individual doing a rapid
test for private reasons are own symptoms but no PCR test, planned weekly leisure meeting
or a symptomatic or positively tested household member}
We assume that there is no private rapid test demand until March when both the citizens'
tests and rapid tests for lay people started to become available
\citep{Bundesanzeiger2021a, Bundesregierung2021} and other access to rapid tests was very
limited.

According to the COSMO study \citep{Betsch2021a} 63\% would have been willing to take a
test in the round of 23rd of February 2021 when an acquaintance would have tested
positive. Since this is only asking for willingness not actual behavior, we take this as
the upper bound of private rapid test demand which we estimate in our model to be reached
in the beginning of May. To cover that many people are likely to have sought and done
their first rapid test before the Easter holidays we add another point that we estimate
for the rapid test demand around Easter. Similarly, we estimate one point in mid March
when tests started to become available in grocery stores and pharmacies which we estimate
in our model. The resulting share of private rapid test demand is shown as the green line
in Figure~\ref{fig:rapid_test_demand} (also see Section~\ref{subsec:estimated_params} for
details on the estimation).

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/testing/rapid_test_demand/baseline_shares}
    \caption{Share of Individuals Doing a Rapid Test.}
    \floatfoot{\noindent \textit{Note:} Rapid test demand can be triggered by individuals
    planning to have education contacts ($\pi_{teacher,\:t}$ or $\pi_{students,\:t}$),
    work contacts ($\pi_{w,\:d}$ and $\pi_{w,\:s,\:t}$), developing symptoms without
    access to a PCR test, having a household member with a positive test or symptoms
    ($\pi_{private,\:t}$). In each case whether a rapid test is done depends on how long
    it has been since the individual's last rapid test and her individual compliance
    parameter. As an example, take a worker in May. In that time workers are encouraged
    to test themselves twice weekly (or every three days, i.e. $\theta_{May,\:work} = 3$)
    but there is no general requirement to test themselves. If the worker has not done a
    test within the last four days in our model she will demand a test if her
    (time-constant) compliance parameter belongs to the upper 60\% in the population.}
    \label{fig:rapid_test_demand}
\end{figure}

\FloatBarrier