### 2022-03-09 src/rev1_reply_comments.tex
Consistent wrapping.

\documentclass[a4paper,11pt]{article}
\usepackage{a4wide}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{afterpage,rotating,natbib,graphicx,setspace,xcolor}
\usepackage{longtable,booktabs,tabularx}
\usepackage{chngcntr}
\usepackage{eurosym,calc}
\usepackage{amsmath,amssymb,amsfonts,amsthm,delarray}
\usepackage{bm}
\usepackage{bbm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xfrac}
\usepackage{listings}
\usepackage{adjustbox}
\usepackage{xr}
\usepackage{xr-hyper}
\usepackage[unicode=true]{hyperref}
\usepackage{afterpage}
\usepackage{clipboard}

\definecolor{darkblue}{rgb}{0, 0, 0.8}
\hypersetup{ colorlinks=true, linkcolor=darkblue, anchorcolor=darkblue,
    citecolor=darkblue, filecolor=darkblue, menucolor=darkblue, runcolor=darkblue,
    urlcolor=darkblue }
\usepackage{enumerate}
\usepackage{caption}
\usepackage{booktabs,caption,rotating}
\usepackage[capposition=top]{floatrow}
% \usepackage{bbm}
\usepackage{epstopdf}
\usepackage{longtable, tabularx, array}
\usepackage{colortbl}

% \usepackage[british]{babel} \usepackage[useregional]{datetime2}
% \DTMlangsetup[en-GB]{showdayofmonth=false}

\usepackage{subcaption}
\usepackage{eurosym}

% \setlength{\marginparwidth}{2cm}
\usepackage[ colorinlistoftodos, backgroundcolor=white, prependcaption, textsize=tiny,
    disable, ]{todonotes}

%%%%PROPOSITION AND PROOF ENVIRONMENTS%%%%
\theoremstyle{plain}
\newtheorem{result}{Result}
\newtheorem{assumption}{Assumption}

\DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max}

% don't let footnotes split across pages
\interfootnotelinepenalty=10000 \widowpenalty=10000 \clubpenalty=10000

\hyphenation{bet-ween}
\hyphenation{with-out}

\frenchspacing

\newlength{\parindentaux}
\setlength{\parindentaux}{\parindent}

\usepackage{enumitem}
\setlist[itemize]{%
    leftmargin=\parindentaux, listparindent=\parindentaux }

% \usepackage{calc}  % Needed to be able to calculate, e.g., sum of dimensions
%\usepackage{ifthen}

\makeatletter
\newcommand*\@myenumerate{enumerate} \newcommand*\@myitemize{itemize}
\newcommand{\comment}[1]{{\color{black!60}#1}} \newcommand{\response}[1]{{%
		\color{black}%
		\par\frenchspacing%
		\ifx\@currenvir\@myenumerate%
		\addtolength{\leftskip}{-\parindentaux}%
		\rm\noindent #1\par%
		\addtolength{\leftskip}{\parindentaux}%
		\else%
		\ifx\@currenvir\@myitemize%
		\addtolength{\leftskip}{-\parindentaux}%
		\rm\noindent #1\par%
		\addtolength{\leftskip}{\parindentaux}%
		\else%
		\rm\noindent #1\par%
		\fi%
		\fi%
}}
\makeatother

\renewcommand{\theequation}{\roman{equation}} \renewcommand{\thetable}{\Roman{table}}
\renewcommand{\thefigure}{\Roman{figure}} \renewcommand{\thesection}{\Roman{section}}

\setlength{\parindent}{0ex}

\begin{document}

\title{\large Reply to the Editor's and Reviewers' comments on: \\[2ex]
    \LARGE The Effectiveness of Antigen Rapid Diagnostic Tests, Vaccinations, and
    Contact Restrictions for Containing the Covid-19 Pandemic\\[2ex]
    \large Submitted to Scientific Reports (Submission ID
    db917d3c-b56a-4129-8eb1-a62a716aaede)\\[-6ex]
}

\author{}
\date{}

\maketitle

\section*{Editors}

\comment{ The title should have a clear, precise scientific meaning and should not
contain a colon. Where possible, the title should be read as one concise sentence. }
\vskip1ex \response{%
    We changed the title from ``The Effectiveness of Strategies to Contain SARS-CoV-2:
    Testing, Vaccinations, and NPIs'' to ``The Effectiveness of Antigen Rapid Diagnostic
    Tests, Vaccinations, and Contact Restrictions for Containing the Covid-19 Pandemic'' }

\section*{Reviewer 1}
\comment{ The authors of the manuscript presented here have used mathematical models to
examine the impact of different measures to contain the SARS-CoV-2 pandemic. Using
epidemiological data from Germany, the contribution of rapid antigen testing is
considered in particular.

The topic studied here is of great importance for the control of the coronavirus
pandemic. The presentation of the scientific methodology is sufficiently detailed,
especially with the supplementary materials. It allows the reader to follow the research
process in detail. However, I noticed a significant deficiency when reviewing the
submitted work. Therefore, the manuscript should definitely be revised again before
publication: A crucial point of the authors is to point out the special benefit of a
rapid test strategy. However, for the calculation of the corresponding influencing
factors in their mathematical model, the authors simplify, specifically, the sensitivity
of rapid antigen tests in an unacceptable way. In the development of the mathematical
model, apparently only the knowledge available a few months ago on the performance of
rapid antigen tests was taken into account. This is evident from literature references
mentioned above. In part, incorrect references are also used, especially for the
assumption of the high sensitivity of the rapid tests. For example, reference No. 74
explicitly does not refer to rapid antigen tests, but to an immunological laboratory
test method (Quidel SARS Sofia antigen FIA). Especially in light of the benefit of rapid
antigen tests described by the authors, it should be considered that many of the
commercially marketed tests show only extremely poor performance. This has recently been
demonstrated by several publications:
\begin{itemize}
    \item     Comparative sensitivity evaluation for 122 CE-marked rapid diagnostic
          tests for SARS-CoV-2 antigen, Germany, September 2020 to April 2021, Scheiblauer,
          Heinrich and Filomena, Angela and Nitsche, Andreas and Puyskens, Andreas and Corman,
          Victor M and Drosten, Christian and Zwirglmaier, Karin and Lange, Constanze and
          Emmerich, Petra and Müller, Michael and Knauer, Olivia and Nübling, C Micha,
          Eurosurveillance, 26, 2100441 (2021),
          \url{https://doi.org/10.2807/1560-7917.ES.2021.26.44.2100441}
    \item Assessment of SARS-CoV-2 rapid antigen tests, Özcürümez, Mustafa, Katsounas,
          Antonios, Holdenrieder, Stefan, von Meyer, Alexander, Renz, Harald and Wölfel,
          Roman, Journal of Laboratory Medicine, vol. 45, no. 3, 2021, pp. 143-148.
          \url{https://doi.org/10.1515/labmed-2021-0036}
\end{itemize}

Especially in the case of a mass application of antigen tests, it is therefore to be
expected that products with lowest sensitivity will also be used in the general
population. The authors should at least model this with an alternative calculation and
discuss it accordingly. Such an improvement would increase the importance of this
manuscript enormously.

Furthermore, I recommend including a declaration of potential conflicts of interest in
the manuscript. This may be particularly important in light of the significance of the
findings from the study for the continued operation of industrial companies.

\vskip1ex

\response{Many thanks for these insightful comments. We respond in three parts.}

\begin{enumerate}
    \item In part, incorrect references are also used, especially for the assumption of
          the high sensitivity of the rapid tests. For example, reference No. 74 explicitly
          does not refer to rapid antigen tests, but to an immunological laboratory test
          method (Quidel SARS Sofia antigen FIA).

          \response{%
              We thank the reviewer for spotting this mistake on our part. We removed
              the reference and only use current information as detailed in the next item. }

    \item In the development of the mathematical model, apparently only the knowledge
          available a few months ago on the performance of rapid antigen tests was taken into
          account.

          The authors should at least model this with an alternative calculation and
          discuss it accordingly. Such an improvement would increase the importance of
          this manuscript enormously.

          \response{%
              We agree that the references we used have become outdated in the meantime
              (we submitted the paper in  August 2021). Indeed, the numbers we used
              turned out to be too optimistic in light of newer studies.

              We therefore ran additional simulations with updated and more conservative
              assumptions on the sensitivity of rapid tests (Appendix B.12) and found that
              our main results are remarkably robust to this change. The percentage of the
              decrease in spring 2021 attributed to rapid tests decreases by only 4
              percentage points even in the least favourable scenario. We describe our
              approach for calculating the adjusted sensitivities in the next answer. }
    \item I recommend including a declaration of potential conflicts of interest in the
          manuscript.
          \response{We have added such a declaration. There are no competing interests.}

\end{enumerate}
}

\section{Reviewer 2}

\comment{ This is a comprehensive and important study that highlights the relevance of
testing to stop the spread of the pandemic.

I cannot comment on the validity of the underlaying mathematical analysis, but I would
like to comment on some assumptions made, which I think require further clarification or
discussion.

\begin{enumerate}
    \item They use contact-frequency from pre-pandemic diary data. Is  this truly useful
          if behavior of people is already changed by the fact there is a pandemic?
          \response{%

              Other models also use the Mossong data: OpenABM-Covid-19 model
              (Hinch2020), Bicher2021.

              We only use the pre-pandemic diary data as the starting point. Our
              policies, which include non-mandated individual behavior adjustments, then
              reduce contacts taking the type of contact into account and for the most
              part not randomly but in very specific ways. This allows us to account for
              behavior reactions and policy effects that target different types of
              contacts differently. For example, household contacts continue to take
              place no matter the policy environment. Work from home does not keep
              random individuals at home but every individual that works has a
              time-constant value that determines when he or she will stay home. Which
              share of the population stays home is taken from Google mobility data. In
              the education sector we do not use the Mossong data but use completely
              mechanistic models. This gives us confidence that we can sufficiently
              accurately represent the contact patterns during the pandemic and account
              for changes induced by the pandemic.

              Also, our excellent fit for most age groups gives us confidence that we
              manage to match actual meeting patterns. If we were far off for example
              for the number of work contacts, we would not be able to match either the
              infections of individuals in the working age (35-59) or of those that
              mostly do not work anymore (60-79).

              \textbf{Answer in case there is confusion because the other multiplier is 1 in fall 2020:}

              We account for individual responses in behaviour to the pandemic through a
              time-variant multiplier that reduces the number of other (i.e. non-work,
              non-household, non-educational) contacts. The values at change points that
              mark policy announcements or fatigue effects are estimated. They are
              detailed in Section A.9.

              Since we estimate both the infection probability of this contact type and
              the multiplier, one of them has to be normalised. We decided to normalise the
              multiplier to 1 in October 2020. This may appear as if we assumed that
              pre-pandemic behaviour had resumed during that time in Germany. However, this
              is not the case. In fact, we would achieve the exact same fit if we doubled
              the infection probability and halved the multiplier. This is because halving
              the multiplier means only half the number of contacts take place. However,
              because of the doubling of the infection probability, in expectation we would
              arrive at the same number of infections in that contact category. }
    \item They state that susceptibility is dependent on age. This requires references
          and needs to be further explained. Also, younger populations are highly susceptible,
          they may just not contribute transmission as much. On that same note, infectiousness
          may not be independent of age

          Edward Goldstein, Marc Lipsitch, Muge Cevik, On the Effect of Age on the
          Transmission of SARS-CoV-2 in Households, Schools, and the Community, The
          Journal of Infectious Diseases, Volume 223, Issue 3, 1 February 2021, Pages
          362–369, \url{https://doi.org/10.1093/infdis/jiaa691}

          \response{%
              We now cite our main reference for the susceptibilities for each age group
              also in the main text. We directly take the susceptibilities given by
              Davies et al. 2020, in their Extended Data Fig. 4. We base our assumption
              on infectiousness being independent of age on Jones et al. 2021. They
              report only slight differences in viral loads between German adults and
              children: ``The least infectious youngest children have 78\% (61, 94) of
              the peak culture probability of adults aged 45 to 55''. Furthermore, they
              caution that these lower numbers may be due to different sample taking -
              such as the usage of smaller pediatric swab devices - rather than actual
              lower viral loads and conclude that the differences, if existent, are
              likely not clinically relevant. }
    \item It is not accurate that the vaccines stop transmission. This has to be adapted
          accordingly. Also, the authors should state which vaccines they based this
          analysis on, as in which ones were in use in Germany at that time and how does
          that affect the relative transmissibility as it cannot be assumed that all
          vaccines will have the same effect. \response{%
              We agree that vaccines do not stop transmission completely. We simplify
              the way vaccines reduce transmission risk considerably by assuming that
              75\% of individuals achieve sterile immunity through a vaccination while
              the remaining 25\% remain fully susceptible. We are comfortable with this
              assumption for several reasons: Firstly, the existing evidence supports
              the assumption that all four vaccines approved in the EU (Biontech/Pfizer,
              Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
              protection in the first few months after vaccination and against the Alpha
              variant (for details see the last third of appendix A.1), which is the
              only time period and variant our simulations cover. Secondly, given the
              surprisingly good performance of rapid tests we tried to err on the side
              of being too optimistic with our modelling of vaccine performance. This is
              why we chose the perfect protection from vaccination for a large part of
              the population. Thirdly, reinfections are not a big issue for the time
              period that we look at. The incidences were small and the chance of more
              than one exposure to Covid are negligibly small. Thus, the value added by
              a probabilistic model of vaccination protection is very small for the
              period that we cover. }
\end{enumerate}
}

% \setstretch{1} \bibliographystyle{ecta} \bibliography{refs}

\end{document}