### 2022-03-10 src/paper/results/robustness_to_sensitivity_assumptions.tex
Mostly done except for Smith reference, switching machines.

\subsection{Robustness to assumptions about rapid test sensitivity}
\label{subsec:robustness_rapid_test_sensitivity}

\Copy{r.1.b-appendix}{
    Our main results are based on rapid test sensitivities read from clinical trials. Recent studies showing that the actual sensitivity of rapid tests may be lower than that \citep[e.g.,][]{Scheiblauer2021}.

    This section shows that our results are robust to making less favorable assumptions
    on rapid test sensitivity. We proceed by describing several possible ways of
    calibrating rapid test sensitivity profiles based on recent studies. Since none of
    these methods is inherently better than the others, we make simulations with two
    sensitivity profiles: The average over all methods and the lower envelope over all
    methods.

    Both profiles imply lower sensitivities of rapid tests than used in our main
    results. This is especially true during the later stage of an infection. However,
    the main results stay very robust. The original result was that rapid tests,
    seasonality and vaccinations are responsible for 42\%, 43\% and 16\%, respectively.
    With the average profile, the effect of rapid tests decreases to 41 \%. With the
    lower envelope profile, which is an extremely unfavorable assumption, it becomes
    38\%.

    % what do we need
    The effect of rapid tests on infection dynamics strongly depends on \emph{when} an
    infection is detected. Earlier detection means that it is more likely that the
    infection has not yet been discovered for a different reason (e.g. due to the onset
    of symptoms) and that the infected person can be isolated before spreading the
    disease to others.

    The sensitivity of rapid tests depends on the viral load in the respiratory tract.
    It is low at the beginning of an infection (especially before the onset of
    infectiousness), high in the first few days of infectiousness and then gradually
    decreasing towards the end of infectiousness.

    We thus need to calibrate a profile of rapid test sensitivities based on the number
    of days until or since the onset of infectiousness.

    % what is available -> two step mapping
    Unfortunately, such sensitivity profiles are not usually reported in studies. We
    thus need to create them by combining two types of studies: 1. Studies that report
    the viral load in terms of threshold cycle ($Ct$) values determined by PCR test
    \citep[e.g.][]{Cosentino2022,Ong2021,Bonnet2022,Jang2021,Zuin2021}. 2. Studies that
    report the sensitivity of rapid tests for different $Ct$ values
    \citep[e.g.][]{Scheiblauer2021,Bruemmer2021}.

    % what options do we have to configure the mappings
    It is natural to assume that the evolution of $Ct$ values over time as well as the
    effect of $Ct$ values on rapid test sensitivity are continuous functions. However,
    the results of the available studies are usually reported in a discretized way. This
    leads to multiple ways of calculating the sensitivity profiles. Some try to recover
    the underlying continuous functions using interpolation or regression, others simply
    use the discretized values.

    % Describe CT value formula vs. bin averages
    For the calibration of $Ct$ values over time we can either use discretized values
    for several time bins from \cite{Ong2021} and \cite{Jang2021}. Alternatively, we can
    use linearized formulas for calculating sensitivities over time \cite{Cosentino2022}
    and complement it with interpolations of data points from \cite{Jang2021} in the
    pre-infectious stage. Throughout we assume that the $Ct$ values of individuals who
    eventually develop symptoms and those who do not follow the same trajectory. This is
    in line with results by \cite{Zuin2021} and recent evidence that rapid tests excel
    at discovering asymptomatic cases \cite{Rosella2022}.

    % Describe sensitivity lookupd vs. interpolation
    For the mapping of $Ct$ values to rapid test sensitivities, again we have two
    options. First, we can simply look up the discretized values for the three $Ct$
    provided in \cite{Scheiblauer2021} (below 20, 25 to 30 and above 35). Secondly, we
    can use linear regression to estimate a continous mapping for the relationship by
    assuming that the $Ct$ values of each bin are achieved exactly at the bin midpoints
    and that the relationship is linear.

    % Properties of the different methods
    In general, using discretized values can lead to an underestimation of peak
    sensitivities and an overestimation of very low sensitivities. This is because
    discretization is essentially a smoothing device. On the other hand, it has the
    advantage of simply working with published results, without introducing any tuning
    parameters or other assumptions.

    % resulting profiles.
    \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/results/figures/data/testing/sensitivity_params_with_different_methods}
        \caption{Rapid test sensitivity profiles}
        \floatfoot{\noindent \textit{Note:} The figure shows estimated sensitivities of
            rapid tests over the course of an infection. The x-axis shows days relative
            to the onset of infectiousness. The y-axis shows the estimated rapid test
            sensitivities. The grey lines are the raw sensitivity estimates obtained
            with different methods of dealing with discretized data. The blue line shows
            their average and the yellow line their lower envelope. The turquoise line
            are the test sensitivities used for the main results of the paper.}
        \label{fig:sensitivity_assumptions}
    \end{figure}

    % \FloatBarrier

    Figure \ref{fig:sensitivity_assumptions} shows that the updated sensitivity
    estimates are lower than the ones used for our original results, especially towards
    the end of an infection. However, the main results barely change. This is due to the
    fact that the differences are most important towards the later stage of an
    infection. Uncovering an infection that was previously undetected at that stage does
    not have a large effect on infection dynamics.

    \begin{figure}
        \centering

        \begin{subfigure}[b]{0.475\textwidth}
            \centering
            \includegraphics[width=0.9\textwidth]{figures/results/figures/data/testing/full_decomposition_channels_area_average}
            \caption{Shapley decomposition with average of sensitivity estimates}
        \end{subfigure}
        \begin{subfigure}[b]{0.475\textwidth}
            \centering
            \includegraphics[width=0.9\textwidth]{figures/results/figures/data/testing/full_decomposition_channels_area_lower_envelope}
            \caption{Shapley decomposition with lower envelope of sensitivity estimates}
        \end{subfigure}

        \caption{Shapley decompositions for different values of rapid test sensitivity}
        \floatfoot{\noindent \textit{Note:} The figure shows updated versions of the
            Shapley decomposition in figure \ref{fig:2021_scenarios_decomposition}. The
            decompositions are based on 20 model runs with different random seeds. The
            share attributed to each channel is rounded to the next full percentage
            point to acknowledge the remaining sampling uncertainty.}
        \label{fig:robustness_to_sensitivity_assumptions}
    \end{figure}

}
\FloatBarrier