### 2022-03-20 src/panel_figures/school_workplace_scenarios.tex

% \documentclass[varwidth,10pt, border=10pt]{wlscirep}
\documentclass[varwidth, border=10pt, class=wlscirep, 10pt]{standalone}

\RequirePackage{helvet}
\RequirePackage{courier}

% border={10pt 5pt} % left/right bottom/top
% border={10pt 10pt 0pt 10pt} % left right bottom top

\usepackage{subcaption}
\usepackage[
    font=footnotesize,
    labelfont={bf,sf,footnotesize},
    labelsep=period,
    justification=raggedright
]{caption}
\usepackage{graphicx}

\begin{document}

\begin{figure}
    \centering

    \begin{subfigure}[b]{6cm}
        \centering
        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/school_scenarios/full_newly_infected}
        \caption{{Effects of different schooling scenarios}}
        \label{fig:school_scenarios}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{6cm}
        \centering
        \includegraphics[width=5cm]{../figures/results/figures/scenario_comparisons/new_work_scenarios/full_newly_infected}
        \caption{{Effects of different work scenarios}}
        \label{fig:workplace_scenarios}
    \end{subfigure}
\end{figure}

\end{document}